### 2020-02-06 02-Chp2-Systematic-Review.Rmd

---
#########################################
# options for knitting a single chapter #
#########################################
output:
  bookdown::html_document2: default
  bookdown::word_document2: default
  bookdown::pdf_document2:
    template: templates/brief_template.tex
documentclass: book
---

```{block type='savequote', include=knitr::is_latex_output(), quote_author='(ref:darwin-quote)'}
There is grandeur in this view of life, with its several powers, having been originally breathed into a few forms or into one; and that, whilst this planet has gone cycling on according to the fixed law of gravity, from so simple a beginning endless forms most beautiful and most wonderful have been, and are being, evolved.
```
(ref:darwin-quote) --- Charles Darwin [@Darwin1859]

# Chapter 2: Systematic Review {#chapter-2-headings}
\minitoc <!-- this will include a mini table of contents-->

## Test