### 2020-02-07 _00-introduction.Rmd
Add word count

---
#########################################
# options for knitting a single chapter #
#########################################
output:
  bookdown::html_document2: default
  bookdown::word_document2: default
  bookdown::pdf_document2:
    template: templates/brief_template.tex
documentclass: book
---

# Covering material {-}

\adjustmtc 
<!-- For PDF output, we must include this LaTeX command after unnumbered headings, otherwise the numbers in the mini table of contents will be incorrect -->

```{r word-count, echo=FALSE, warning=FALSE}
words <-
  wordcountaddin::word_count("01-Chp1-Introduction.Rmd") +
  wordcountaddin::word_count("02-Chp2-Systematic-Review.Rmd")

date <- as.character(format(Sys.time(),"%d%m%Y"))

tmpWords <- data.frame(words = words, date = date)

masterWords <- read.csv("words/words.csv", header = TRUE)

masterWords <- rbind(masterWords,tmpWords)

write.csv(file = "words/words.csv", x = masterWords, col.names = FALSE, row.names = FALSE)
```

Word count: `r words`