### 2020-11-25 01-Chp1-Introduction.Rmd
This time for sure

---
#########################################
# options for knitting a single chapter #
#########################################
output:
  bookdown::html_document2: default
  bookdown::word_document2: default
  bookdown::pdf_document2:
    template: templates/brief_template.tex
documentclass: book
---

# Introduction {#intro-intro}
\minitoc <!-- this will include a mini table of contents-->

<!--
The {#rmd-basics} text after the chapter declaration will allow us to link throughout the document back to the beginning of this chapter. Reference labels will automatically be generated if you don't specify them, as the lowercase title with spaces replaced by hypens (e.g. r-markdown-basics). Look for the reference to this label at the beginning of the next chapter.
-->

## Aims and Objectives of the thesis

### Aim

```{r}
f <- flextable::flextable(robvis::data_rob2)

f <- flextable::autofit(f)

f
```

### Objective

## Chapter Overview

* __Chapter \@ref(background-heading):__ Background information on dementia and blood lipid levels. This chapter provides an introduction to the topics covered in this thesis to non-subject area experts, and discusses the motivation for the remainder of the thesis. 
* __Chapter \@ref(sys-rev-tools-heading)__ This Chapter introduces new evidence synthesis two tools built in R: `robvis` and `medrixvr`. The motivation for, development and impact of, and future plans for these tools are outlined.
* __Chapter \@ref(sys-rev-heading)__ This Chapter describes the
* __Chapter \@ref(cprd-analysis):__ This Chapter examines the relationship between lipid-regaulting agent use and dementia outcomes in the Clinical Practice Research Datalink, a large primary care electronic health record database, based in England. 
* __Chapter \@ref():__ This Chapter seeks

## Thesis Output

The outputs of this thesis are detailed below, and include published peer review papers, presentations at conferences, open source evidence synthesis tools, and . . .

### Peer reviewed papers

__Include other papers I have been involved in?__

### Papers under review

* McGuinness et al.

### Software

* `robvis`: An R package and associated `shiny` web application that allows users to easily visualise the results of risk of bias graphs.

* `medrxvir`: An R package and associated `shiny` web application that allows users to easily search and retrieve bibliographic data from the medRxiv^[https://www.medrxiv.org/] preprint repository.

### Talks

* Talk at Cochrane on medrxivr
* Webinar to Evidence Synthesis Ireland on Risk of Bias test

## Summary

*
* 
*