### 2020-11-30 front-and-back-matter/abbreviations.tex

% do not edit by hand - add to abbreviations.csv instead
% First parameter can be changed eg to "Glossary" or something.
% Second parameter is the max length of bold terms.
\begin{mclistof}{List of Abbreviations}{3.2cm}
\item[DOI] Digitial object identifier
\item[HDL] High Density Lipoprotein
\item[MR] Mendelian randomization
\end{mclistof}