### 2021-04-08 front-and-back-matter/abbreviations.tex

% do not edit by hand - add to abbreviations.csv instead
% First parameter can be changed eg to "Glossary" or something.
% Second parameter is the max length of bold terms.
\begin{mclistof}{List of Abbreviations}{3.2cm}
\item[API] Application programming interface
\item[AzD] Alzheimer's disease
\item[CIND] Cognitive impairment not dementia
\item[CPRD] Clinical Practice Research Datalink
\item[CRAN] Comprehensive R Archive Network
\item[DOI] Digitial object identifier
\item[HDL] High density dipoprotein
\item[IPD] Individual participant data
\item[LDL] Low density lipoprotein
\item[MCI] Mild cognitive impairment
\item[MMSE] Mini Mental State Exam
\item[MoCA] Montreal Cognitive Assessement
\item[MR] Mendelian randomization
\item[NOS] Newcastle-Ottowa Scale
\item[PDF] Portable document format
\item[RCT] Randomised controlled trial
\item[TG] Triglycerides
\item[VaD]  Vascular dementia
\end{mclistof}