### 2021-04-08 tmp.Rmd

Chapter 1. Introduction
1.1. Summary
This chapter presents the problem that the research conducted in this thesis seeks to
address and an outline of the main output of this thesis: new evidence concerning
repurposing antihypertensives for dementia prevention using genetic and non-genetic
instrumental variable analyses. This chapter then introduces three issues that arose and
have been addressed during the conduct of this research. They are: how to calculate
power for instrumental variable analyses in pharmacoepidemiology; how to use
Mendelian randomization to predict drug repurposing opportunities; and what factors
have effected prescribing of existing dementia drugs. This chapter concludes with the
aims and objectives of this thesis, a summary of its organization and a list of the
associated outputs.
1.2. Statement of the problem
There is a substantial unmet clinical need for treatments for dementia where significant
benefits to patients, society and the public purse can be gained. Despite this, some drug
companies have recently withdrawn from this therapy area due to failed and costly efforts
to find new treatments. (1,2) Drug repurposing, the identification of properties in existing
or abandoned compounds for other clinical conditions, offers significant advantages over
traditional drug discovery approaches. This includes immediate access to human safety
data from the original clinical development work, which can accelerate testing in clinical
trials, saving both time and money. (3–5)
Many antihypertensive medications have been proposed as drug repurposing candidates
for the prevention of dementia. In part, because of research to better understand several
reports of observed associations between midlife hypertension and later-life risk of
Alzheimer’s disease and vascular dementia. (5–8) There is also increasing recognition
that one of the earliest pathological events in the development of Alzheimer’s disease is
vascular dysregulation. (9) As well as suggestions that some antihypertensives,
specifically those that block angiotensin receptor and calcium channel signalling, may
have other neurological benefits. (9–11)
2
Several observational studies have investigated repurposing antihypertensives for
dementia prevention. (12–19) However, these studies have typically used case-control
designs with logistic regression and cohort designs with survival analysis, both of which
are observational study designs that may be subject to unmeasured or residual
confounding and reverse causation. Specific concerns of observational studies are
confounding by indication, where the reasons that a patient receives a treatment relate to
the reasons that the patient is at an increased risk of the outcome; and healthy adherer
bias, where patients initiating or adhering to a drug for prevention of a condition are
more likely to be healthy. There is also potential for reverse causation due to preclinical
or early stages of the disease, which could lead to more frequent contacts with general
practitioners (GPs) and lead to raised blood pressure being more likely to be detected.
This, in turn, could lead to the prescription of an antihypertensive drug in advance of
dementia being formally diagnosed. Consequently, current evidence concerning
repurposing antihypertensives for dementia prevention is considered inconclusive.
1.3. New evidence concerning repurposing antihypertensives for dementia prevention
using instrumental variable analysis
Instrumental variable analysis, which estimates the causal effect of an exposure on an
outcome by using a third variable (the instrument), can be robust to confounding and
reverse causation if certain assumptions are met. That is:
  IV1. The instrument must associate with the exposure
IV2. The instrument must only affect the outcome through the exposure
IV3. The instrument and the outcome must have no common causes
These assumptions can be represented on a directed acyclic graph, as shown in Figure
1.1. Instrumental variable analysis, with other sources of evidence, can be used in a
triangulation framework to obtain a reliable answer concerning the potential repurposing
of antihypertensives for dementia prevention. (20) This thesis presents two forms of
instrumental variable analysis to provide new evidence concerning this hypothesis. The
first uses physicians’ prescribing preference as a non-genetic instrument in electronic
health record data obtained from the United Kingdom (UK) Clinical Practice Research
Datalink (CPRD). The second uses single nucleotide polymorphisms (SNPs) i.e.
differences in the deoxyribonucleic acid (DNA) nucleotides between individuals, which
have been selected to mimic the biological function of the protein targets of 
3
antihypertensive drug classes, as a genetic instrument in an approach more commonly
known as Mendelian randomization. The results of both these instrumental variable
analyses have been made available via the following references. (21,22)
1.4. Calculating power for instrumental variable analysis in pharmacoepidemiology
Instrumental variable analysis is an increasingly popular method in the field of
pharmacoepidemiology. (23–28) However, the power calculators that were available for
studies using instrumental variable analysis at the start of my PhD – such as Mendelian
randomisation power calculators – did not allow for the structure of research questions
using non-genetic instruments (for example, physicians’ prescribing preference) in this
field. (29,30) This is because analysis using non-genetic instruments in pharmacoepidemiology will typically have stronger instruments and so can detect smaller causal
effects. Consequently, there was a need for dedicated power calculators for these type of
research questions in pharmacoepidemiology. In this thesis, I investigate how to conduct
Figure 1.1: A directed acyclic graph illustrating the basic instrumental variable analysis model.
Directed acyclic graphs are a visual representation of a model, which represent variables by ‘nodes’
and the relationships between them by ‘directed edges’. (68) The graphs are defined as acyclic
because edges cannot form ‘cycles’, whereby the edges all act in the same direction. This prevents a
variable from being both a cause and a consequence of itself. Directed acyclic graphs are a common
tool in epidemiology and, more specifically, causal inference when edges are given a causal
interpretation, as they allow researchers to depict their model and its assumptions using a common
graphing scheme. (69)
4
power calculations for pharmacoepidemiological studies, which use a single binary
instrument to analyse the causal effect of a binary exposure on a continuous outcome. I
also provide an online calculator, as well as packages in both R and Stata, for the
implementation of the formula by others (https://github.com/venexia/PharmIV). This
work has been published in the International Journal of Epidemiology. (31)
1.5. Using Mendelian randomization to predict drug repurposing opportunities
Identification of drug repurposing opportunities can maximize the benefit of a drug.
However, as highlighted before, the more traditional observational research methods
used to investigate these opportunities are subject to several biases. These include
confounding by indication, reverse causality, and missing data. In this thesis, I propose
Mendelian randomization as a novel approach that can be used for the prediction of drug
repurposing opportunities. Mendelian randomization addresses some of the limitations
associated with the existing methods in this field. Furthermore, it can be applied either
pre- or post-approval of the drug and could therefore prevent the potentially harmful
exposure of patients in clinical trials and beyond. This thesis includes discussion of
examples from the literature that have used Mendelian randomization to predict drug
repurposing opportunities and covers the strengths and limitations associated with using
this method for this purpose. There was relatively little discussion focussed on using
Mendelian randomization for drug repurposing when I commenced my PhD and I have
since published on this topic. (32)
1.6. Factors effecting existing dementia drug prescribing
Drugs for dementia have been available in England from 1997. However, since their
launch, there have been several changes to national guidelines and initiatives that may
have influenced prescribing. These include changes in National Institute for Health and
Care Excellence (NICE) guidance; several government dementia strategies; the addition
of dementia to the Quality and Outcomes Framework (QOF); and the expiry of drug
patents. Despite this, little research had been conducted prior to my PhD into the effect
of these events on prescribing. (33,34) In this thesis, I investigate prescribing trends in
England since the launch of these drugs up to 1st January 2016 using data from the
CPRD to address this gap in the literature. The key motivation for this analysis was to 
5
identify factors that have affected prescriptions of existing treatments that may also
influence repurposed drug candidates in the future. However, the results from this
analysis could also be used to identify breaks in the prescription of these drugs that could
be exploited as natural experiments for progression studies. For example, if these drugs
were not prescribed when the NICE guidelines stopped recommending their use between
2006 and 2011, the progression of people diagnosed during this time who did not access
the drugs could be compared with the progression of people diagnosed before and after
this time who did have access to the drugs. The results presented in this chapter have
been published in Alzheimer’s Research & Therapy. (35)
1.7. Aims and objectives
The objective of this thesis was to use instrumental variable analysis methods, in existing
data sources, to triangulate evidence for repurposing antihypertensive drugs for the
prevention of dementia.
The specific aims were as follows:
  1. Develop a power calculator for non-genetic instrumental variable analysis studies
in the context of pharmacoepidemiology.
2. Describe the use of genetic instrumental variable analysis, namely Mendelian
randomization, for predicting drug repurposing opportunities.
3. Examine the impact of regulatory guidance and patent expiry on dementia drug
prescribing.
4. Investigate whether antihypertensive drugs have a causal effect on incident
dementia using instrumental variable analysis with electronic health record data.
5. Investigate whether antihypertensive drugs have a causal effect on incident
dementia using instrumental variable analysis with genetic data.
1.8. Organization of this thesis
Chapter 2 provides the background necessary for the rest of this thesis through the
introduction of dementia, the concept of drug repurposing, and discussion of the existing
evidence regarding antihypertensive drugs for dementia prevention. It also covers the 
6
strengths and limitations of observational pharmacoepidemiology and explains why it
might be preferable over ‘gold standard’ randomized controlled trials for some
hypotheses. Chapter 3 describes instrumental variable analysis, the method utilized
throughout this thesis, and documents the development of a power calculator for this
method in the context of pharmacoepidemiology (Aim 1). Chapter 4 introduces the idea
of using Mendelian randomization, a form of instrumental variable analysis that uses
genetic variants as instruments, for drug repurposing and covers the strengths and
limitations of this approach (Aim 2). Chapter 5 describes the CPRD and my use of this
data source. Chapter 6 covers the current treatments available for dementia and factors
affecting their prescription in England based on data from the CPRD (Aim 3). Chapter 7
presents an assessment of the effects of antihypertensive drugs on dementia prevention
using instrumental variable analysis with data from the CPRD (Aim 4); while Chapter 8
presents an assessment of the effects of antihypertensive drugs on dementia prevention
using instrumental variable analysis with genetic data (Aim 5). The thesis concludes with
a discussion in Chapter 9 that brings together all the elements of this thesis and discusses
their implications.
1.9. Outputs from this thesis
1.9.1. Contributions to scientific literature
Contributions to scientific literature arising from this thesis are detailed below and
provided in Appendix A.
The protocol for the observational work using CPRD data to investigate drug
repurposing opportunities is published in the BMJ Open. (36) Its contents are referenced
in several places throughout this thesis, particularly Chapter 5:
  Walker VM, Davies NM, Jones T, Kehoe PG, Martin RM. Can commonly prescribed
drugs be repurposed for the prevention or treatment of Alzheimer’s and other
neurodegenerative diseases? Protocol for an observational cohort study in the UK
Clinical Practice Research Datalink. BMJ Open. 2016 Dec 1;6(12):e012044.
7
The power calculator for instrumental variable analysis in pharmacoepidemiology,
described in Chapter 3, is published in the International Journal of Epidemiology (31):
  Walker VM, Davies NM, Windmeijer F, Burgess S, Martin RM. Power calculator for
instrumental variable analysis in pharmacoepidemiology. Int J Epidemiol. 2017 Oct
1;46(5):1627–32.
Also published in the International Journal of Epidemiology is an article discussing
Mendelian randomization as a novel approach for the prediction of adverse drug events
and drug repurposing opportunities (32), which formed the basis of Chapter 4:
  Walker VM, Davey Smith G, Davies NM, Martin RM. Mendelian randomization: a
novel approach for the prediction of adverse drug events and drug repurposing
opportunities. Int J Epidemiol. 2017 Dec 1;46(6):2078–89.
The trend analysis examining prescribing practice for drugs for dementia in the CPRD,
presented in Chapter 6, is available from Alzheimer’s Research & Therapy (35):
  Walker VM, Davies NM, Kehoe PG, Martin RM. What is the impact of regulatory
guidance and expiry of drug patents on dementia drug prescriptions in England? A trend
analysis in the Clinical Practice Research Datalink. Alzheimer’s Research & Therapy.
2018 May 29;10:51.
The assessment of antihypertensives for dementia prevention using electronic health
record data, reported in Chapter 7, is currently under peer review and available from
bioRxiv (21):
  Walker VM, Davies NM, Martin RM, Kehoe PG. Comparison of antihypertensive drug
classes for dementia prevention. bioRxiv. 2019 Jan 12;517482.
The assessment of antihypertensives for dementia prevention using genetic data, reported
in Chapter 8, has been made available as an advance article from the International
Journal of Epidemiology (22):
  8
Walker VM, Kehoe PG, Martin RM, Davies NM. Repurposing antihypertensive drugs
for the prevention of Alzheimer’s disease: a Mendelian Randomization study. Int J
Epidemiol. 2019 Jul 4; Advance article.
1.9.2. Contributions to scientific meetings
I presented the paper “Power calculator for instrumental variable analysis in
pharmacoepidemiology”, described in Chapter 3, at the UK Administrative Data
Research Network Annual Research Conference 2017 in Edinburgh, UK.
I presented the paper “Mendelian randomization: a novel approach for the prediction of
adverse drug events and drug repurposing opportunities”, which formed the basis of
Chapter 4, at the University of Bristol Population Health Symposium 2016 in Bristol, UK
and at the International Society for Pharmacoepidemiology mid-year meeting 2018 in
Toronto, Canada.
I presented posters, based on the paper “What is the impact of regulatory guidance and
expiry of drug patents on dementia drug prescriptions in England?” detailed in Chapter
6, at the Alzheimer’s Research UK Research Conference 2016 in Aberdeen, UK and at
the University of Bristol brain research showcase and networking day 2018 in Bristol,
UK.
I presented “Can treatments for hypertension be repurposed for the treatment of
dementia?” at the Society of Epidemiologic Research annual conference 2018 in
Baltimore, United States of America (USA); at the European Congress of Epidemiology
2018 in Lyon, France and the International Society for Pharmacoepidemiology annual
meeting 2018 in Prague, Czech Republic. These presentations combined results from
Chapters 7 and 8.