### 2021-04-12 tmp.Rmd
Test

---
output: html_document
bibliography: bib/library.bib
csl: csl/nature.csl
space_betwee_paragraphs: true
fig_caption: true
always_allow_html: yes
link-citations: true
---

```{r include=FALSE}
# word count = 
doc.type <- knitr::opts_knit$get('rmarkdown.pandoc.to') # this will tell the .Rmd file what output you are knitting to (word, pdf, html) so that you cna use if else statements when making tables - html/pdf output tables do not work well in word. For word we need to use the 'flextable' packaged to make tables.
## packages
library(dplyr)
library(kableExtra)
library(flextable)
```

# Data visualisation {#chapter6}

<style>
body {
text-align: justify}
</style>

### *EpiViz: a tool to visualise large metabolomic association analyses* {-}

In Chapters \@ref(chapter4) and \@ref(chapter5) we observed large metabolic shifts as a result of increased adiposity. These results were presented using Circos plots. Production of the first stage of these plots was inefficient. This chapter details the development and implementation of a web application and `R` package that efficiently creates Circos plots that can be used by the community for similar analyses. The `R` package was used to re-make the plots presented in Chapters \@ref(chapter4) and \@ref(chapter5) and make those in subsequent chapters. <br>

\newpage

## Introduction
The metabolome (briefly discussed in Section \@ref(metabolites)) is a complex biological system comprised of the bodies entire small molecules. These small molecules, metabolites, take many forms and act in a multitude of pathways[@Griffin2006]. Metabolites are expressed differentially among cell types and carry information from genetic and non-genetic factors[@Johnson2016; @Wishart2019]. They are the ultimate end-point of a biological process and are therefore seen as the link between genotype and phenotype[@Griffin2006; @Su2014; @Chu2019; @Wishart2019]. <br>

There are potentially many thousands of metabolites[@Griffin2006] across different cell types, with expression associated with traits and diseases[@Guasch-Ferre2016; @Liesenfeld2013; @Johnson2016; @Moore2014; @Cirulli2019; @Wishart2018]. However, assigning individual metabolite changes to diseases is difficult[@Su2014; @Fearnley2016; @Chu2019] because of the complexity of their interrelationships. The metabolome is made up of many feedback and feed-forward loops, meaning metabolic products of one pathway will likely influence those of another. The intercorrelated nature of metabolites[@Rosato2018] is further complicated by the wide degree of genetic correlation[@Gallois2019] and shared genetic architecture[@Shin2014; @Kettunen2016; @Long2017; @Gallois2019]. Gaining global overview of association analyses will likely aid interpretation of results. Global overview is of particular importance in MR studies where robust selection of instrumental variables is required to meet methodological assumptions[@DaveySmith2003]. <br>

Forest plots are widely used in metabolomic association analyses[@May-Wilson2017; @Taylor2019; @Soares2019; @SantosFerreira2019; @Hartley2020; @Borges2017] to visualise and interpret results. They can be easily made, with a number of `R` packages, such as [`ggforestplot`](https://github.com/nightingalehealth/ggforestplot) and [`TwoSampleMR`](https://mrcieu.github.io/TwoSampleMR/#to-many-forest-plot)[@Hemani2018], available. However, when presenting more than 50 data points forest plots are cumbersome and many separate plots are needed. This makes gaining global overview difficult. <br>

An alternative approach is to compress information into a circular rather than vertical/horizontal form (see Figures 3 and 2 from [@Shin2014] and [@Kettunen2016] respectively). Circos plots[@Krzywinski2009], as implemented in many genetics studies, condense large amounts of information into informative visualisations (see Figure 1A from [@Saben2014]). However, Circos software is designed for genomic data and written in programming languages unfamiliar to many epidemiologists. The `R` package [`circlize`](https://jokergoo.github.io/circlize_book/book/index.html)[@Gu2014] provides most of the functionality of the Circos software but is designed for genomic information and requires an advanced understanding of `R`. Here I present `EpiViz`, a web application and `R` package, that builds on the [`circlize`](https://jokergoo.github.io/circlize_book/book/index.html)[@Gu2014] and [`ComplexHeatmap`](https://jokergoo.github.io/ComplexHeatmap-reference/book/)[@Gu2016] `R` packages. `EpiViz` enables epidemiologists with little programming experience to efficiently create Circos plots with hundreds of variables and gain global overview.

## Methods
### Circos plots for association analyses
`EpiViz` adapts and builds on the `Circlize` foundation to create Circos plots compatible with association analysis data. Circos plots are composed of six elements: template, plotting space, tracks, sections, data, legend (Figure \@ref(fig:annotated-circos-plot)). The template element is a square of defined proportions within which information is plotted. Each additional element is layered onto the template one after the other. The plotting space element is an empty circle which is layered and centred on top of the template. Data is plotted on to the plotting space. An optional extra of the Circos plot, the legend element takes the dimensions of the template and creates a separate plotting space that can be layered on to the bottom of the template element. <br>

The plotting space is separated into tracks and sections. Tracks are laid down as rings within the plotting space. Each track represents a single element of information such as an exposure. Tracks are numbered from the outside to the centre of the circle and coloured separately. Sections divide the plotting space into distinct areas, much like a pie chart. Sections are defined by the data and usually represent groups of outcomes such as metabolite classes. A section track is laid at the outside of the tracks to give a header for each section. The header is referenced in the legend element. <br>

Once the template, plotting space, tracks and sections are laid down, coordinates for each section and track location can be called to plot the data element. Each track and section coordinate, e.g. track 2 section 3, is treated as an individual plotting space. As such, data can be plotted based on the following coordinates: track, section, $X$, $Y$. The $X$ axis of each track is defined by the number of rows in the data frame, i.e a data frame with 100 rows will have an $X$ axis of length 100 with each row given an $X$ axis coordinate from 1:100. The $Y$ axis is defined by the minimum and maximum of the data for that track. As such, each track and section coordinate, e.g. track 2 section 3, can be considered an individual plot with a $Y$ axis that is shared by all of the sections in that track. For each position on the $X$ axis the label element of each row is plotted outside of the section header. <br>

```{r annotated-circos-plot, echo=FALSE, out.width='100%', fig.cap="Annotated Circos plot highlighting elements of the Circos plot"}
knitr::include_graphics("data/chapter6/figures/annotated_circos_plot.pdf")
```

### Implementation
`EpiViz` is a `Shiny` web application and `R` package. `R` (version 3.6.2)[@r2019] and `Shiny`[@Chang2019] (version 1.4.0) were used to develop the web application and `R` (version 3.6.2) was used to develop the `R` package. `Shiny` is an `R` package that enables development and deployment of web applications written in the `R` programming language. Development of `EpiViz` was progressive and feedback from colleagues (see Section \@ref(epiviz-acknowledgements)) was vital in this process.

### Operation
The [web application](http://mrcieu.mrsoftware.org/EpiViz/) is publicly accessible -- [http://mrcieu.mrsoftware.org/EpiViz/](http://mrcieu.mrsoftware.org/EpiViz/) -- and held under an [MIT license](https://github.com/mattlee821/EpiViz/blob/master/LICENSE.txt). [Docker](https://www.docker.com/) was used to containerise the application. The [Medical Research Council Integrative Epidemiology Unit](http://mrcieu.mrsoftware.org/EpiViz/) at the University of Bristol `Shiny` server hosts the web application. The web application has been tested on computers running macOS (version 10.14) and Windows (version 10) using: Internet Explorer (version 11; Windows), Google Chrome (version 79; macOS and Windows), Safari (version 13; macOS). <br>

The [`R` package](https://github.com/mattlee821/EpiViz/tree/master/R_package) is publicly accessible through GitHub -- https://github.com/mattlee821/EpiViz/tree/master/R_package --  and held under an[MIT license](https://github.com/mattlee821/EpiViz/blob/master/LICENSE.txt). The `R` package is accessible on all computers with `R` version 3.3.0 or higher and has been tested on macOS (version 10.14) and Windows (version 10) running `R` version 3.3.0 or higher. <br>

A legend function is available for both the web application and `R` package and is implemented using functions from the [`ComplexHeatmap`](https://jokergoo.github.io/ComplexHeatmap-reference/book/)[@Gu2016] `R` package. By default the colours used for the Circos plot in both the web-application and `R` package are accessible colours identified using [i want hue](https://medialab.github.io/iwanthue/). Example data is provided on the web application *Home* tab and with the `R` package using the function `EpiViz::EpiViz_data*()` where `*` is 1-3. Example data can be produced for use with the web-application and `R` package using code on [GitHub](https://github.com/mattlee821/EpiViz/blob/master/R_package/data-raw/example_data_simulation.R). <br>

#### `Shiny` web application
In order to use the [web application](http://mrcieu.mrsoftware.org/EpiViz/) a web-browser and an internet connection of at least 1Mbps is required. No other system requirements are needed. Upon opening the web application, users are shown a number of example Circos plots created with the application and are directed towards the [*Home*]() tab. The *Home* tab (Figure \@ref(fig:web-app-home)) provides users with a short summary of the application, link to the `R` package, and example data for use with the app. <br>

```{r web-app-home, echo=FALSE, out.width='100%', fig.cap="Screenshot of the EpiViz web application Home page"}
knitr::include_graphics("data/chapter6/figures/web_app_home.png")
```

The [*How to*]() provides instructions for using the application and is split into three: Step 1 deals with the preparation of data to be used with the application, step 2 deals with how to use the application, and step 3 provides information on potential next steps. Users are instructed to upload one data frame per track of the Circos plot. Each data frame should be a tab delimited text file and `R` code is provided for users to achieve this. The user is guided through an example utilising the example data provided with the application and downloadable on the *Home* tab. <br>

Having prepared their data as instructed in Step 1, users select the [*Analysis*]() tab and upload one data frame for each track of the Circos plot. If formatted as described in the *How to* tab, descriptive information including a volcano plot will be produced automatically (Figure \@ref(fig:web-app-analysis-step1)). <br>

```{r web-app-analysis-step1, echo=FALSE, out.width='100%', fig.cap="Screenshot of the EpiViz web application Analysis page, showing Step 1"}
knitr::include_graphics("data/chapter6/figures/web_app_analysis_data.png")
```

If the data frame appears as expected to the user the Circos plot can then be produced by selecting the [*Plot*]() tab. A series of drop down lists are presented which auto-populate with the column names from the first uploaded data frame. The following column information is required for the Circos plot: label, group, estimate, p-value, lower confidence interval, upper confidence interval (Figure \@ref(fig:web-app-analysis-step2)). Users can select one, two, or three tracks for their Circos plot, with data for the respective track coming from the upload data section. <br>

```{r web-app-analysis-step2, echo=FALSE, out.width='100%', fig.cap="Screenshot of the EpiViz web application Analysis page, showing Step 2"}
knitr::include_graphics("data/chapter6/figures/web_app_analysis_plot.png")
```

A p-value threshold (default 0.05) can be provided to indicate a significance threshold. On the Circos plot a solid point is indicated as reaching the significance threshold and an open point indicates not reaching the threshold. An optional legend function is provided and users can choose the labels for the legend points. The legend is auto-populated using the levels in the `group` column of the up-loaded data frame. Finally, users can select to use *Accessible colours* (default) or *Not accessible colours* for the colours of the points on the Circos plot. <br>

#### `R` package
The `EpiViz` `R` package is accessed using `R` version 3.3.0 or above. Documentation for using the package to create Circos plots is available as a README on [GitHub](https://github.com/mattlee821/EpiViz/tree/master/R_package) and includes use cases and troubleshooting. The `R` package can be installed directly from GitHub into `R` with the following `R` code: <br>

```{r eval=FALSE}
# Install devtools
install.packages("devtools")
library(devtools)

# Install directly from GitHub
devtools::install_github("mattlee821/EpiViz/R_package")
library(EpiViz)
```

Once installed, users can use the example data provided with the package to produce Figure \@ref(fig:r-package-3-track-type-plot), which illustrates the three different types of track available (point, line, bar), using the following `R` code: <br>

```{r eval=FALSE}
circos_plot(track_number = 3,
            track1_data = EpiViz::EpiViz_data1,
            track2_data = EpiViz::EpiViz_data2,
            track3_data = EpiViz::EpiViz_data3,
            track1_type = "points",
            track2_type = "lines",
            track3_type = "bar",
            label_column = 1,
            section_column = 9,
            estimate_column = 2,
            pvalue_column = 3,
            pvalue_adjustment = 1,
            lower_ci = 4,
            upper_ci = 5,
            lines_column = 2,
            lines_type = "o",
            bar_column = 2)
```

```{r r-package-3-track-type-plot, echo=FALSE, out.width='100%', fig.cap="Circos plot prduced using EpiViz R package and example data"}
knitr::include_graphics("data/chapter6/figures/r_package_3_track_type_plot.pdf")
```

These Circos plots are designed to be viewed on a screen rather than on paper. As such the plots are created at a larger scale compared to traditional plots made in `R`. This means that when viewing the Circos plot in an viewer such as the `R` Studio Viewer pane the plot appears squashed. Viewers like this should be used only as a guide when making the plot. Instead, plots should saved as `PDF` using the below code -- `PDF` files can be converted to other image formats. <br>

```{r eval=FALSE}
pdf("my_circos_plot.pdf",
    width = 30, height = 30, pointsize = 35)
circos_plot(...
            circle_size = 25)
dev.off()
```

`R` sets default height, width and pointsize parameters and users should iterate the sizing of the plot adjusting the `width` and `height` functions to get the desired plot size. They should then adjust the `pointsize` function to increase and decrease the size of all information in the plot (labels, points, lines etc.). The numbers provided work with most plots and were used for all examples. In addition, users can adjust the size of the Circos plot directly using the argument `circle_size`. The default for `circle_size` is 25 and smaller numbers increase the size of the circle and larger numbers decrease the size. Finally, users can adjust the height of tracks individually using `track*_height` where `*` refers to the track number (see below code). The default for each track is 20 percent of the total size of the circle (the section track is fixed at 5 percent). <br>

```{r eval=FALSE}
circos_plot(...
            track1_height = 0.2,
            track2_height = 0.2,
            track3_height = 0.2)
```

In order to minimise the time required to maintain the `R` package further customisation is achieved by the users themselves. The function is written to aid this customisation with *Default parameters* and *Customisable parameters* located at the top of the `circos_plot()` function script. Guidance and code is provided on [GitHub](https://github.com/mattlee821/EpiViz/tree/master/R_package) to aid user customisation, for example changing track colours requires users navigating to the *Customisable parameters* section and then to the *Colours* section and replacing the following lines with their desired colours: <br>

```{r eval=FALSE}
discrete_palette <- c("#00378f", # track 1 colour
                      "#ffc067", # track 2 colour
                      "#894300") # track 3 colour
```

## Discussion
Circos plots have been used to visualise and provide overview of large association analyses both in the literature[@Shin2014; @Kettunen2016] and this thesis (see Chapters \@ref(chapter4), \@ref(chapter5) and \@ref(chapter9)). This has been time consuming and cumbersome. Requiring bespoke code these plots have been challenging to work with when using multiple data sets (see [GitHub](https://github.com/mattlee821/000_thesis/tree/master/index/data/chapter6/code/circos_example.R) for example bespoke code). <br>

`EpiViz` provides an efficient way to produce Circos plots for association analyses and can be implemented with little up-front cost. The web application is intended for researchers with limited to no experience of `R`, while those with some experience can gain added functionality from the `R` package. The application of Circos plots to comparing and gaining global overview of association analyses is presented in Chapters \@ref(chapter4), \@ref(chapter5) and \@ref(chapter9), as well as by colleagues[@Taylor2019]. <br>

Though intended to be low maintenance, there is scope for further development of both the web application and `R` package. Of broad appeal to both is the inclusion of new plotting methods such as the chord diagram which shows relatedness between two nodes. In the case of metabolites a chord diagram would provide greater understanding about the relationship between different sections. In addition, the ability to filter and choose specific sections to display would simplify the plot and highlight potentially meaningful results. <br>

The web application is less flexible than the `R` package, however there are specific extensions of interest. Firstly interactive plots would provide users the ability to expand sections of interest and display additional information. Rather than a final presentation of the data this would primarily focus on exploration of data. Additionally, introducing features from the `R` package such as different plot types will broaden the appeal of the web application. These additions are likely to increase the costs of maintenance however. <br>

The `R` package has extensive scope for enhancement and presents a more efficient investment of time for extensions over the web application due to the flexibility of `R`. Currently, only a single data point can be plotted on each track at each $X$ axis coordinate. This is to prevent over-crowding. However, if a data set has 50 points for example, adding a second point to each $X$ axis coordinate of the track would not overcrowd the plot. <br>

The current limited customisation afforded with the web application and the `R` package is designed to reduce the maintenance costs of the author. As a result, functions that users may wish to use, such as multiple point estimates on a single track, must be coded by the user in `R`. Though guidance is provided on GitHub for such customisations, this additional effort may reduce the utility of the tool, especially among users with little experience of `R` but very specific requirements. Similarly, although designed to be efficient, users must provide data in a specific format in order to use the web application and `R` package. `R` code is provided to aid this effort but the requirement for users with no `R` experience to create a specific data frame will be off-putting for some. <br>

In order to achieve the desired goal of providing global overview of large association analyses Circos plots must be much larger than traditional plots such as forest plots. As such, Circos plots are best utilised as a tool for the user rather than a reader of a print journal. Circos plots are best viewed using a screen with the ability to zoom in to sections where more detail is required. This fact will reduce the utility of this type of Circos plot in publications as fixed margins will limit the size and resolution of the plot. Resolution can be improved by saving the Circos plots as a `PDF` file. Many journals do not accept `PDF` images and will convert these into lossless formats such as `PNG` or `JPEG`. This will reduce the resolution provided for readers. <br>

`EpiViz` is simple and efficient to use. There is potential for `EpiViz` to become widely used among the community. To pre-empt popularity and provide a stable environment the web application was containerized using Docker and hosted on a server with provision for multiple-users. With popularity it is hoped that a community will help to develop `EpiViz` further. <br>

## Acknowledgements {#epiviz-acknowledgements}
`EpiViz` was greatly aided by the `R` packages [`Circlize`](https://jokergoo.github.io/circlize_book/book/index.html)[@Gu2014] and [`ComplexHeatmap`](https://jokergoo.github.io/ComplexHeatmap-reference/book/)[@Gu2016], whose code is freely available and was used as a reference point. The following contributed to the development of `EpiViz`: Luke McGuinness, Osama Mahmoud, Sam Neaves. The following colleagues provided feedback in the development of `EpiViz`: Caroline Bull, Charlie Hatcher, Kurt Taylor, Nancy Mcbride, Neil Goulding, Steph Suddell.

\newpage