### 2021-04-15 02-Chp-Background.Rmd
End of day: -2625

---
bibliography: bibliography/references.bib
csl: bibliography/nature.csl
output:
  bookdown::pdf_document2:
    template: templates/brief_template.tex
  bookdown::word_document2: default
  bookdown::html_document2: default
documentclass: book
---

```{block type='savequote', include=knitr::is_latex_output(), quote_author='(ref:dara-quote)', echo = TRUE}
Science knows it doesn't know everything; otherwise, it'd stop.
```

(ref:dara-quote) --- Dara O'Briain

```{r, echo = FALSE, warning=FALSE, message=FALSE}
source("R/doc_options.R")
source("R/helper.R")
knitr::read_chunk("R/background.R")
doc_type <- knitr::opts_knit$get('rmarkdown.pandoc.to') # Info on knitting format
```

# Background, Theoretical framework, Aims & Objectives {#background-heading}

\minitoc <!-- this will include a mini table of contents-->

__Do I need another section on why the relationship is unclear?__

## Figures/Tables To

* Diagnostic criteria
* Statin mechanism of action 
* Evidence pyramid
* Overview of MR studies

## Lay summary

Around 850,000 people in the UK live with dementia, and by 2040, nearly twice as many will have the condition. Despite many promising candidates, no cure for dementia currently exists, meaning the focus is on finding ways to prevent the condition. The best way to do this is to find risk factors (characteristics that influence a person's chance of developing a disease) for dementia that we can easily change. Avoiding a risk factor does not guarantee that a person will not develop dementia, but makes it less likely. A key risk factor for dementia may be the levels of lipids (fatty substances such as cholesterol) in a person's blood, though not all existing research agrees. The aim of this thesis is to use all available evidence to assess whether blood lipids levels are in fact a risk factor for dementia.

This introductory chapter provides background information on both dementia and blood lipids, and on the potential link between them. It introduces the theory used to frame the research presented here, and then maps the formal aims and objectives of the research project to the relevant chapters of this thesis. Finally, it summarises the outputs (journal articles, presentations and software) that were created as part of this thesis.

<!----------------------------------------------------------------------------->

## Introduction

This chapter provides an overview of the broad context of this thesis, introducing the core concepts used throughout and providing some background on each. It briefly discussed the underlying pathologies and diagnosis of dementia, its public health importance, and the current state of treatment and prevention therapeutics.

This chapter then provides background on blood lipids and lipid-modifying treatments, and summarises the existing evidence on the effect of these exposures on dementia outcomes.

It introduces evidence synthesis as the key framework used to guide the research presented in the remaining chapters. I will outline the aims, objectives and structure of this thesis, and

Finally, it briefly summarises the contributions to the scientific literature that arose from this research.

<!----------------------------------------------------------------------------->

## Dementia

Dementia is major neurocognitive disorder, with symptoms including impairment of executive cognitive functions such as speech, judgement and memory.

<!----------------------------------------------------------------------------->

### Underlying pathologies {#underlying-pathologies}

Dementia is a collective umbrella term for a range of conditions each with different clinical presentations and disease courses, and is generally __defined as ""__. <!--- Cite --->

Alzheimer's disease, named after Dr. Alois Alzheimer who presented the first recorded case in 1907, is the most common cause of dementia, accounting for approximately 60-80% of cases. Characterised by substantial cognitive impairment and diffculty with high level executive function to the extent that it interferes with, it is an insidious disease, within initial onset thought to occur up to 15 years prior to symptomatic presentation. Despite research implicating the accumulation of amyloid plaques and neurofibrillary tangles, or alternatively the cholingeric nervous system, <!--- check ---> as potential mechanisms of disease, much remains unknown unknown about Alzheimer's pathogenesi

<!-- [__Need a bit more here__] -->

Vascular dementia (VaD) is the second largest underlying pathology of dementia, accounting for ~10% of cases and is caused by a range of cerebrovascular disorders. Presentation can vary widely based on the underlying cause. Due to the varied underlying  pathophysiology, VaD can either onset quite rapidly following a cerebrovascular event such as a stroke or over a long time-frame due to a series of small infarcts. VaD is diagnosed using the __NINCDS-AIREN__ criteria.[@roman1993vascular]

The remaining 10-30% of cases are caused other dementia subtypes (e.g. Lewy Body dementia) or by progression of other neurological diseases (e.g. Parkinson's disease).

The term mixed dementia is used when more than one underlying dementia pathologies co-occur. This is a  common presentation of dementia because the presence  of one underlying condition  weakens overall brain health and increases the likelihood  of another condition developing. <!--- Double check this --->

<!-- __MCI__ and difference between MCI and dementia, MCI as a stepping stone to dementia, etc. -->

<!----------------------------------------------------------------------------->

### Diagnostic criteria {#diagnostic-criteria}

Dementia is difficult to diagnose, primarily due to the absence of a gold standard test for the condition. <!-- is this true, or would it be better to say that there is no gold standard test without an autopsy?--> There exist several diagnostic criteria for dementia and related disease. In all cases, dementia is diagnosed on the basis of behavioural and cognitive changes as assessed by an experience clinician. 
<!-- This means the potential for misclassification both of dementia versus no-dementia and between the different underlying conditions is extremely likely. -->

<!----------------------------------------------------------------------------->

```{r diagnosticCriteria-setup, results="hide", message=FALSE, echo = FALSE}
```

```{r diagnosticCriteria-table, message=FALSE, results="asis", echo = FALSE}
```

<!----------------------------------------------------------------------------->

Two of the most commonly used criteria include those found in the Diagnostic and Statistical Manual of Mental Disorders, fourth edition (DSM-IV-TR) <!-- __[CITATION NEEDED]__ --> criteria and the National Institute of Neurological Disorders and those developed by the Stroke-Alzheimer's Disease and Related Disorders Association (NINCDS-ADRDA).[@dubois2007]

<!-- [**Need first NINCDS-ADRDA citation and DSM-IV citations**] -->

While these define the overall parameters for a assinging a dementia diagnosis, they are informed using practical scales to measure patients cognitive health, with two of the best known of these scales being the Mini Mental State Exam (MMSE) and Montreal Cognitive Assessment (MoCA) scale.

<!-- Also touch on aspects of each, such as the clock-drawing and whether or not they need to validated with input from another person. -->

<!-- Note important differences between the scales and whether they affect results. Could cross-reference with an exploratory analysis in systematic review -->

The distinction between these memory scales and diagnostic criteria presented above should be noted. For example, the MMSE is a tool for assessing cognitive impairment, but does not by itself indicate the absence or presence of dementia, instead merely indicating congitive impairment which could be due to another cause (for example, temporary delerium as a result of an infection or surgery). <!--- Check --->
 A low score on the MMSE should feed into the diagnostic process, potentially as a screening tool to help identify which patients require further assessment by a trained clinician. Studies that use a memory scale alone are likely to be biased.

<!-- Differentiating between the underlying causes of a dementia diagnosis is challenging but necessary, as whether the patient has Alzheimer's disease or vascular dementia will affect expect progression and potential treatment options available. -->

<!-- ### Differing risk by sex -->

<!-- Transparency guidelines that suggest sex should eb treated as biological variable and so results should be presented stratified t -->

<!-- Rate of dementia/AD differs in men and women <!--- Though is this just due to the different lifestyles and life expectancies --->

<!-- Previous studies have hinted that the relationship is dependent on individual-level characteristics such as age and sex [16,28]. -->

<!----------------------------------------------------------------------------->

<!-- ## Challenges in the study of dementia using traditional study designs -->

<!-- Speak to the -->

<!-- Historically, studies on dementia have faced a range of challenges. As a range of determinants (genetic, environmental and lifestyle) are thought to jointly influence the risk, progression and outcomes of dementia, any individual association is likely to have only a small effect. Therefore, the statistical power need to detect these associations will often require sample sizes that are unfeasible using primary data collection. This fact is further complicated by the long latency of dementia, which necessitates a costly long-term approach to patient follow-up. Furthermore, dementia studies often have limited generalisability to the target population, as certain subgroups (e.g. the very old) are frequently under-represented in study cohorts due to difficulties associated with their recruitment.9 Finally, studies employing a case-control design are further limited by the high potential for differential recall bias between those who have and have not developed dementia.10 -->

<!-- Fortunately, these limitations may potentially be addressed through the use of routinely collected health data, and the electronic health record databases in which they are stored. -->

<!-- *Cognitive impairment not dementia* (CIND) is another term used to described those with cognitive impairment but who fall below the diagnostic criteria for dementia. -->

<!-- A further clinical subtype that is of particular import the -->

<!-- And MCI -->

<!-- hr -->

<!-- I -->

<!-- Consider sex as a biological variable -->

<!----------------------------------------------------------------------------->

### Public health importance

Dementia is quickly becoming a critically important public health issue. Despite the age-specific incidence and prevalence of dementia remaining relatively constant over time,[@prince2016] an ageing population is set to create a dementia epidemic, particularly in Westernised countries. While approximately 525,000 patients have received a dementia diagnosis, the true number of people currently living with dementia in the UK is thought to be closer to 850,000, with this figure expected to double by 2040.[@baker2019] Globally, the prevalence of dementia is expected to reach 75 million by 2030.[@prince2016]

Dementia also has a substantial economic impact. In 2015, the estimated total cost of dementia in England was £24.2 billion. Health care costs alone were £3.8 billion.[@wittenberg2019] <!--- Cross check against old data on cost to NHS --->. Based on the implications of a dementia diagnosis, this is perhaps unsurprising. Patients living with dementia are more prone to falls and other conditions that necessitate admission to hospital. Once admitted, patients with dementia are more likely to be stay in hospital for longer.[@mollers2019]

<!-- For the same period, unpaid care was estimated at .[@wittenberg2019] -->
<!-- While the burden of dementia on patients themselves... comment here on the economic strain of unpaid care and the emotional strain of lookign after someone who has dementia - part. related to emotional and personality changes. -->

The urgent need to reduce the burden of dementia, both at the personal and system (national health systemic) level, is clear. As such, the systematic assessment of easily modifiable targets (such as blood lipid levels) for their utility in the prevention of dementia should be prioritized.

<!----------------------------------------------------------------------------->

<!-- ### Risk factors -->

<!-- Dyslipidemia (see Section \@ref(evidence-association) for a fuller dicussion). -->

### Treatments

Developing treatments for dementia is regularly deemed to be one of the hardest markets in the pharmaceutical world, with trials of seemingly promising therapeutics, such as BACE inhibitors, being regularly abandoned due to futility.[@cummings2020]. At present, there are no known curative treatments for dementia, regardless of the underlying cause, though several therapeutics can help alleviate the symptoms of Alzheimer's disease.

The most common of these are acetylcholinesterase (ACE) inhibitors, which inhibit the degradation of the neurotransmitter acetylcholine by competitively binding the ACE enzyme. Acetylcholine plays a key role in controlling the cholingeric synapses, which are highly concentrated in regions of the brain (such as the neocortex) that control higher level brain functions such as memory and attention.[@hampel2018] Commonly prescribed ACE inhibitors include donepezil and galantamine.[@pariente2008] ACE inhibitors increase the availability of the neurotransmitter, and has shown clinical effect is easing the behavioural and memeory-related symptoms of Alzheimer's disease.[@marucci2020] ACE inhibitors are only a stop-gap treatment, treating the symptoms rather than the underlying pathology which may continue to progress.

### Risk factors

Given the substantial burden that dementia represents and the absence of any curative therapies, a substantial amount of research has been produced examining putative risk factors for dementia.[@feingold2000;@peters2019] The most commonly studied are cardiovascular factors, including hypertension,[@hughes2020association], stroke[@kuzma2018] and hyperlipidaemia. 
In this context, lipid levels represent a promising target for preventative treatment, due to the ready availability of lipid-modifying treatments which could be repurposed.[@pushpakom2019a] Determining whether variations in lipid levels are causative for dementia may prove critical in reducing the future burden of the condition.

The next section provides an overview of blood lipid fractions and therapeutic interventions that modify them, while Section \@ref(evidence-association) provides an overview of the evidence for an association between lipids and dementia outcomes.

<!----------------------------------------------------------------------------->

## Serum lipids

<!----------------------------------------------------------------------------->

### Lipid fractions

The blood lipid profile contains a range of fractions. However, for the sake of this thesis, we will only consider the two most important fractions, trigylcerides (TG) and cholesterol, which are either absorbed from food (exogenous lipids) or produced internally (endogenous lipids).[@feingold2000]

Triglycerides are the simplest and most common type of lipids found across the body. They are used to store unused calories from food, and to move energy around the body. In contrast, cholesterol is primarily used to create cell walls and certain sex hormones. As lipids are not water soluble, within the blood stream, cholesterol is transported in lipoprotein structures of varying densities. Low-Density-Lipoprotein-cholesterol (LDL-c), commonly know as the "bad" cholesterol, transports fat to cells. In contrast, High Density-Lipoprotein-cholestel (HDL-c), transports cholesterol to the liver to be broken down and excreted.

<!-- Need to also touch on the interplay between the different subcategories of lipids Does lowering one fraction affect the other(s)? Are there feedback loops? This is important because if not, statins can only be used to triangulate with findings for LDL-c. May also need to standardise in terms of -->

<!--- Will need to explain why I am focusing on these fractions rather that looking at others include VLDL, etc --->

<!----------------------------------------------------------------------------->

<!-- Measurement error is a big deal for this thesis, as will discuss measurement error regarding how  different codelists/measurement of AD vs vascular dementia affect results in later chapters. -->

Total serum cholesterol (TC) is a commonly-used summary measure to estimate the total amount of lipid present in the blood. The measure is derived from measurements of the individual HDL-c, LDL-c and TG levels using the Friedwald formula:[@friedewald1972]

$$TC \approx LDLc - HDLc - kTG$$
where $k$ is 0.20 if measurements are in milligrams per decilitre (_mg/dl_) and 0.45 if measured in millimole per litre (_mmol/l_).

Widely used criteria for assessing the acceptable levels of different types of lipids are based on the National Cholesterol Education Program (NCEP)[@national2002third], and are outlined in Table \@ref(tab:lipidLevels-table).

&nbsp;

<!----------------------------------------------------------------------------->

(ref:lipidLevels-caption) Classification of blood lipid levels according to the National Cholesterol Education Program guidelines.[@national2002third]

(ref:lipidLevels-scaption) Classification of blood lipid levels

```{r lipidLevels-table, message=FALSE, results="asis", echo = FALSE}
```
<!----------------------------------------------------------------------------->

&nbsp;

Elevated cholesterol in the bloodstream via high LDL-c levels leads to atherosclerosis, the build-up of fatty deposits in the blood vessels. These deposits constrict blood flow and can lead to vascular complications. Alternatively, part of the deposit can detach from the artery walls, forming a clot that can lead to a heart attach or stroke. Globally, the prevalence of elevated total cholesterol, or hypercholesterolemia, was estimated by the World Health Organisation to be approximately 40%.

<!----------------------------------------------------------------------------->

<!-- ### Hypercholesterolemia -->

<!-- Clinical diagnosisof cholesterol above a certain level - moves from a continuous . realtively improtant in EHR where blood lipid levels may not be measured, but the formal diag -->

<!-- Scales for Hypercholesterolemia in UK - see NICE guidance -->

<!-- Can also be familial Hypercholesterolemia - -->

### Statins

Statins are by far the most common method of lipid regulation currently prescribed. Statins inhibit the conversion of 3-Hydroxy-3-Methylglutaryl-Coenzyme-A (HMG-CoA) into mevalonic acid,  by competitively binding with HMG-CoA Reductase (HMG-CoA-R). This conversion limits the rate at which cholesterol (see Figure \ref(), can be produced, enabling statins to effectively reduce the production of LDL cholesterol.

<!----------------------------------------------------------------------------->

(ref:statin-mechanisam-cap) **Statin mechanism of action:** Overview of the

(ref:statin-mechanisam-scap) Statin mechanism of action

```{r statin-mechanisam, echo = FALSE, results="asis", fig.cap='(ref:statin-mechanisam-cap)', out.width='100%', fig.scap='(ref:statin-mechanisam-scap)'}
knitr::include_graphics(file.path("figures/sys-rev/prismaflow.png"))
```

<!----------------------------------------------------------------------------->

Statins can either be lipophilic or hydrophilic, with several of the common brands falling into each category. The distinction between the two is of particular importance in the assessment of their impact on dementia, as only lipophilic statins can cross the blood brain barrier. However, this is only applicable if the relationship between statins and dementia risk is independent of their impact on circulating blood levels.

### Other lipid regulating agents (LRA)

There are a range of other interventions that can be used to modify a persons lipid profile, which each acting in slightly different ways. However, in general, these treatments are either used as adjunct (additional) treatments with statins therapy or are used in situations where statins are contra-indicated or not tolerated.

<!-- see https://www.heart.org/en/health-topics/cholesterol/prevention-and-treatment-of-high-cholesterol-hyperlipidemia/cholesterol-medications for information on the indications for treatment - could present as a table -->

The most commonly used non-statin therapeutic is ezetimibe, which prevents intestinal absorption of cholesterol. However, when used alone, it has a limited LDL-c lowering effect, leading to the creation of combined statin/ezetimibe therapies (both compounds contained in a single pill, as opposed to complimentary treatments).

A second example are fibrates, used to treat hypertriglyceridaemia by reducing production of triglyceride carrying compounds in the liver. They are commonly used in patients with mixed hyperlipidaemia if treatment with statins has failed to sufficiently control cholesterol levels.

Finally, PCSK9 inhibitors (or PCSK9i) are a relatively new treatment with strong lipid lowering effects, lauded as a potential alternative to statins.[@chaudhary2017] Their mechanism of action is to bind to and inhibit PCSK9, which breaks down LDL-c receptors on the surface of the liver, thus allowing more LDL-c to be internalised and broken down.

&nbsp;

<!----------------------------------------------------------------------------->
(ref:lipidTreatments-caption) Summary of available treatments for hyperlipidaemia.

```{r lipidTreatments-table, message=FALSE, results="asis", echo = FALSE}
```
<!----------------------------------------------------------------------------->

&nbsp;
&nbsp;

<!----------------------------------------------------------------------------->

## Evidence for the association between blood lipids and dementia {#evidence-association}

This section provides an overview of the varying sources of evidence on the relationship between blood lipid levels and dementia risk.

<!----------------------------------------------------------------------------->
### Basic science

The core underlying biological cause of demetnia,

A role for lipids in the aetiology of the disease is supported by both genetic linkage studies and functional cell biology studies.

The generation of the amyloid plaques found in the brains of Alzheimer's patients is cholesterol dependent [CITE], while the most established genetic risk factor for late-onset dementia, apolipoprotein E (ApoE), is involved in cerebral cholesterol transport. Several other genes involved in cholesterol transport have also been found to be associated with increased AD susceptibility [CITE].

<!--- might need some more here --->

Despite these results, evidence from the diverse range of epidemiological studies on this topic has been inconclusive.

<!----------------------------------------------------------------------------->
### Observational studies

Several observational studies have examined the relationships between concentrations of serum lipids (total cholesterol (TC), low density lipoprotein cholesterol (LDL-c), high density lipoprotein cholesterol (HDL-c) and triglycerides) and both Alzheimer's disease and vascular dementia and reported extremely varied results. In some studies, a high serum cholesterol concentration has been found to be associated with an increase in susceptibility to AD [8-13], however others have shown no association [14-17], or a reduced susceptibility [@mielke2005; @reitz2004a]. With regards VaD, decreased levels of HDL-c appear to be associated with increased risk [18,20,21], while for LDL-c, studies have reported both positive and negative association [18,22].

<!--- Will need more here - particularly on the --->

<!-- Need to talk about biases here? Or have a different section: Diverse sources of evidence: different underlying biases -->

<!----------------------------------------------------------------------------->
### Randomised controlled trials

In terms of the central research of this thesis, RCTs of statin therapy can be used to provide indirect evidence for the effect of reducing blood LDL-c levels on dementia risk.

However, RCTs may be infeasible if the outcome of interest is one with a long prodomal period, such as dementia (see Section \@ref(underlying-pathologies)), as they would require extremely long and costly follow-up.[@ritchie2015] It is no surprise then that the two previous trials providing evidence on the effect of statins on dementia risk, identified by a recent Cochrane review,[@mcguinness2016a] are in fact trials of statins for the prevention of coronary related outcomes.

While being widely cited, these studies have major limitations that their utility to provide evidence on the effect of statin treatment on in assessing the impact of lipid-lowering treatment on dementia risk. Firstly, there was no clinical cognitive evaluation of patients to determine a dementia outcome. One of the trials, the Prospective Study of Pravastatin in the Elderly (PROSPER) trial,[@trompet2010] reported not on dementia outcomes but on the change in cognitive scores over a mean of 3.2 years. As highlighted in Section \@ref(diagnostic-criteria), a "change in score" alone is insufficient to diagnose a dementia outcome. The second trial, the Medical Research Council/British Health Foundation Protection Study,[@2002] found no effect of simvastatin on dementia (`r estimate(1, 0.61, 1.65)`), but did not report how the outcome was assessed/recorded within the trial.

Additionally, the two trials did not make any effort to assign an underlying pathology to each case, instead reporting an all-cause dementia outcome. As discussed in Section \@ref(underlying-pathologies), the different underlying pathology of dementia have different mechanisms of action, and so it is not given that the effect of statins would be consistent across them.

Both trials were also limited by the relatively short follow-up period examined, expected when the primary outcome of the trials were coronary related conditions rather than dementia.[@2002; @trompet2010] The PROSPER trial had a mean follow-up of 3.2 years, while the MRC/BHF Protection Study estimated risk at 5 years of follow-up. Given the long lag time between non-symptomatic onset of dementia and clinical presentation. Finally, as they included only patients at high vascular risk, their generalisability to other settings is limited.[@mcguinness2016b]

<!----------------------------------------------------------------------------->

### Mendelian randomisation

Newer methodological approaches,such as Mendelian randomisation (MR),[@daveysmith2014] have also been used to examine the effect of varying lipid levels on dementia risk in an effort to combat the risk of reverse causation and residual confounding inherent to observational studies. In brief, MR uses genetic variants that are both strongly associated with the exposure of interest and are independent from potential confounders to strengthen causal inference.[@daveysmith2014] The analytic method relies on several assumptions about the instrumental variable (IV),[@davies2018] as shown in Figure \@ref(fig:mr-overview), namely that:

1. the IV is associated with the exposure of interest (the relevance assumption);
2. the IV and outcome do not share a common cause (the independence assumption); and
3. the IV does not affect the outcome other than via the exposure (the exclusion restriction assumption).

Several Mendelian randomisations studies have examined the effect of blood lipid levels on dementia outcomes, finding <!--- Finish --->

A recent MR study indicated that low levels of LDL-c may cause a reduction in AD risk. However, this study was widely criticised as it did not exclude the region surrounding the ApoE gene, the strongest known risk factor for Alzheimer's disease. This invalidated the exclusion exclusion restriction criteria (Assumption 3, above), as the risk reduction observed in this study could be driven by variants in this region via a pathway independent of lipid levels.

<!--- Need more on MR here --->

<!----------------------------------------------------------------------------->

&nbsp;

In summary, multiple sources of evidence exist on the relationship between statins and dementia. In the next section, I introduce the theoretical framework used in this thesis.

Multiples forms of evidenc

<!----------------------------------------------------------------------------->

## Theoretical framework: Evidence synthesis

Evidence synthesis is the process of finding and integrating information from several sources to examine a research question.[@donnelly2018a]

The results of an evidence synthesis exercise can be used to provide a more definitive answer to that question or, failing that, to highlight gaps in the existing evidence base. The ability to identify these gaps is particularly useful in guiding future research to address questions that have yet to be answered.

This thesis seeks to use an evidence synthesis framework to assess the effect of lipids, and treatments that influence lipid levels, on dementia outcomes. 
Specifically, this thesis considers three methodologies within the umbrella term of evidence synthesis:

* Triangulation
* Inclusion of grey literature
* Individual patient data meta-analysis

These three elements are expanded on below and are used to frame the research presented in the subsequent Chapters.

<!----------------------------------------------------------------------------->

### Triangulating across study designs

<!----------------------------------------------------------------------------->

As illustrated in Section \@ref(evidence-association), several diverse epidemiological methods have been used to examine the effect of varying blood lipid levels on dementia risk. However, each method is limited by its own unique biases. Aetiological triangulation is a developing evidence synthesis method that seeks to exploit these inherent differences in study design, and as a result, in biases.[@lawlor2016a] If several sources of evidence are available and point towards identical conclusions about a exposure-outcome relationship, and these sources are at risk of unrelated biases, this strengthens our confidence in the result. The ideal scenario is where predicted sources of bias are likely to be in competing directions, one expected to increased the effect of the exposure and the other to attenuate it.[@lawlor2016a] As such, triangulating these results provides us with a middleground, between the directions of bias. A triangulation approach can also prove useful in a prospective manner, helping to design new studies that are at risk of different sources of bias to that already available from the published literature.

In order to triangulate across different evidence sources, all available evidence on a research question must first be identified. This requirement motivates the conduct of the comprehensive systematic review presented in Chapter \@ref(sys-rev-heading).

Due to the fact triangulation exercises revolve around integral biases, it is necessary to assess the direction and magnitude of bias in a study. Risk-of-bias assessment tools provide a structured and transparent way to evaluate the internal validity of a study across several domains of bias. Tools now exist for a range of study designs including:

* randomised controlled trials;[@sterne2019] 
* non-randomised studies of interventions;[@sterne2016robins] and
* diagnostic accuracy studies.[@whiting2011quadas]

Risk-of-bias assessment using a domain-based tool is a recommended part of the systematic review process. [@Pagen71; @sterne2019a; @mcguinness2018] In spite of this, several of the existing highly-cited systematic reviews of observational studies of the relationship between statins and dementia did not perform any risk-of-bias assessment[@chu2018b] or used an outdated tool.[@anstey;@poly2020c] As such, a core component of the systematic review presented in Chapter \@ref(sys-rev-heading) is a comprehensive domain-based risk-of-bias assessment for all included studies.

The findings from the review will then be triangulated with the primary analyses performed in this thesis (Chapters \@ref(cprd-analysis-heading) and \@ref(ipd-analysis-heading)) in Chapter \@ref(dicussion-heading).

<!----------------------------------------------------------------------------->

### Inclusion of grey literature {#diverse-sources-preprints}

The importance of including grey (or gray) literature in systematic reviews is widely acknowledged. Meta-research studies have demonstrated that systematic reviews excluding grey literature sources overestimate the effect of interventions.[@conn2003; @mcauley2000; @hopewell2007] Common, well-accepted forms of grey literature include conference abstracts and theses.[@lefebvre2019searching]

A important developing source of grey literature are preprints. Defined by the Committee on Publication Ethics (COPE) as 'scholarly manuscript[s] posted by the author(s) in an openly accessible platform, usually before or in parallel with the peer review process'[@committeeonpublicationethicscope2018], preprints serve several purposes. They are used to establish primacy when submitting to a journal where the peer-review process may take several months,[@vale2016] to rapidly disseminate research findings, as occurred during the COVID-19 pandemic,[@fraser2020a] and to make available publications that may not have been accepted elsewhere in an attempt to combat publication bias or the "file-drawer" effect.[@rosenthal1979]

One of the major criticisms of using preprints as an evidence source is that they have not yet undergone formal peer review.[@maslove2018; @schalkwyk2020] However, this approach assigns a lot of weight to peer-review as a indicator of "quality", and is at odds with the acceptance of non-reviewed conference proceedings as an evidence source.[@lefebvre2019searching; @mahood2014] The argument for including preprints is further strengthened by evidence that the results from preprinted studies seldom change following peer review. Meta-studies of the concordance between preprinted and published studies showed that results were broadly comparable between the two, indicating that while the numerical results may change, the overall interpretation of the results were consistent in the overwhelming majority of cases.[@shi2021a; @klein2019; @nicholson2021] This indicates that preprints should be considered a reliable reflection of a given study.

In this thesis, preprints are considered an important source of evidence. However, as with many sources of grey literature, there are several logistical issues with carrying out systematic searches in preprint repositories.[@mahood2014] As such, to enable their inclusion in the systematic review described in Chapter \@ref(sys-rev-heading), a new tool addressing these issues is presented in Chapter \@ref(sys-rev-tools).

<!----------------------------------------------------------------------------->

### Individual patient data meta-analysis

If lipids are found to have a causal role in development of dementia, evidence-based preventative strategies would be informed by identifying the types of individuals who are most likely to receive benefit from treatment with lipid-modifying agents.

Analysis of summary-level data, that is the data extracted from publications describing primary studies, can only take investigators so far in this regard. Due to a lack of primary studies readily presenting results stratified by covariates of interests, for example sex and ethnicity (see Section \@ref()), meta-analyses of summary-level data often have limited ability to examine exposure-covariate interactions (for example, sex and ethnicity.

An individual participant data (IPD) meta-analysis is therefore the best option to examine the modification of dementia risk by individual-level covariates. This analytical approach has the added benefit of allowing a common set of inclusion criteria and statistical model to be applied across all dataset, potentially eliminating some important sources of heterogeneity. It also allows for the incorporation of relevant, previously un-analysed datasets, available via new initiatives such as the Dementia Portal UK,<!--- Cite this ---> which aim to provide access to several dementia-related datasets via a single simplified application process.

This thesis will attempt to obtain the raw data from relevant primary studies identified by the systematic review in Chapter \@ref(sys-rev-heading). Any data contained will be combined with previously unanalysed data from the DPUK portal as part of an individual participant data meta-analysis in Chapter \@ref(ipd-analysis-heading).

<!----------------------------------------------------------------------------->

## Thesis overview

### Hypothesis

Circulating blood lipid levels, and by extension treatments that modify blood lipid levels such as statins, affect the risk of subsequent dementia.

### Aims and objectives

The specific research objectives that this thesis seeks to address are:

* To create a tool that allows for the inclusion of health related preprints in evidence syntheses in a systematic and reproducible manner
* To review all available evidence across multiple diverse study designs to assess the effect of lipids and lipid regulating agents on dementia risk
* To examine whether there is evidence for an effect of lipid-regulating agents on dementia and related outcomes in a large scale population-based cohort, the Clinical Practice Research Datalink (CPRD)
* To meta-analyse previously unexplored datasets as part of a individual participant data (IPD meta-analysis)

### Structure

Chapters are self-contained, presenting the methods and results of that specific research project. The are bookended by introductory and discussion sections which place the methods and results in context. Each chapter is prefaced by a "Lay" or plain English summary, developed with input from the Patient and Public Advisory Group (see Section \@ref(disc-PPI) for a discussion of the group's involvement and Appendix \@ref(appendix-ppi) for more detail on the group).

-   **Chapter \@ref(background-heading):** Background information on dementia and blood lipid levels. This chapter provides an introduction to the topics covered in this thesis to non-subject area experts, and discusses the motivation for the remainder of the thesis.
-   **Chapter \@ref(sys-rev-tools-heading):** This Chapter introduces a new tool, `medrxivr`, which was used to developed to allow for systematic searches of the health-related preprint repositories.
-   **Chapter \@ref(sys-rev-heading):** This Chapter describes a comprehensive systematic review and meta-analysis of all available evidence on the relationship between blood lipids, and interventions that modified blood lipids, and dementia.
-   **Chapter \@ref(cprd-analysis-heading):** This Chapter examines the relationship between lipid-regulating agent use and dementia outcomes in the Clinical Practice Research Datalink, a large primary care electronic health record database, based in England.
-   **Chapter \@ref(ipd-heading):** This Chapter describes an individual patient data analysis of several previously unanalysed longitudinal cohort studies, to describe the relationship between blood serum lipids and dementia outcomes.
-   **Chapter \@ref(discussion-heading)**: This Chapter integrates the diverse evidence identified and produced as part of this thesis. The overall strengths and weaknesses of this project are discussed in detail, and further work is suggested.

An overview of how all research studies in this thesis can be found in Table \@ref(tab:thesisOverview-table)

\blandscape
<!----------------------------------------------------------------------------->
(ref:thesisOverview-caption) Summary of studies included in this thesis, and used as evidence sources in the triangulation exercise performed in Chapter \@ref(discussion-heading). Note, Chapter \@ref(sys-rev-tools-heading) is intentionally not included in this table, as it describes a tool rather than a research study.

```{r thesisOverview-table, message=FALSE, results="asis", echo = FALSE}
```
<!----------------------------------------------------------------------------->
\elandscape

## Outputs from this thesis {#thesis-output}

The outputs of this thesis are detailed below, and include published peer review papers, presentations, and open-source evidence synthesis tools.

### Contributions to the scientific literature

During the course of this thesis, I have made several contributions to the scientific literature. Those arising from or directly related to the contents of this submission are presented below.

&nbsp;

_**McGuinness, L. A.**, and L Schmidt. (2020) "medrxivr: Accessing and searching medRxiv and bioRxiv preprint data in R." Journal of Open Source Software 5.54 2651. DOI: [10.21105/joss.02651](https://doi.org/10.21105/joss.02651)_

A paper introducing the open-source preprint search tool described in Chapter \@ref(sys-rev-tools-heading). As is common for journal articles describing software, the paper is intentionally short providing only a broad overview of the tool while extensive documentation is available from the project website (see \@ref(sys-rev-tools-intro) for more details).

&nbsp;

_Hennessy, E. A., Acabchuk, R., Arnold, P. A., Dunn, A. G., Foo, Y. Z., Johnson, B. T., Geange, S. R., Haddaway, N. R., Nakagawa, S., Mapanga, W., Mengersen, K., Page, M., Sánchez-Tójar, A. Welch, V.,  **McGuinness L. A.** (2020). Ensuring Prevention Science Research is Synthesis-Ready for Immediate and Lasting Scientific Impact. MetaArXiv. DOI: [10.31222/osf.io/ptg9j](https://doi.org/10.31222/osf.io/ptg9j)_

The experience of extracting data for the systematic review in Chapter \@ref(sys-rev-heading) inspired a practical guide for researchers in prevention science is currently under review at Prevention Science. This piece was co-written with Dr. Emily Hennessy (see Author Declarations in the front materials).

&nbsp;

_**McGuinness, L. A.**, and Higgins J. P. T. (2020) "Risk‐of‐bias VISualization (robvis): An R package and Shiny web app for visualizing risk‐of‐bias assessments." Research Synthesis Method). DOI: [10.1002/jrsm.1411](https://doi.org/10.1002/jrsm.1411)_

The tool used to visualise the risk-of-bias assessments in Chapter \@ref(sys-rev-heading) has been published in Research Synthesis Methods. See Appendix \@ref(appendix-robvis) for more details on this tool.

&nbsp;

_**McGuinness, L. A.**, and Sheppard A. L. 2020. “A Descriptive Analysis of the Data Availability Statements Accompanying Medrxiv Preprints and a Comparison with Their Published Counterparts.” MetaArXiv. DOI: [10.31222/osf.io/p75xe](https://doi.org/10.31222/osf.io/p75xe)_

Using the tool described in Chapter \@ref(sys-rev-tools), I lead a "research-on-research" study to assess the concordance between the data availability statements accompanying a sample of medRxiv preprints and their published counterparts.

&nbsp;

For information on additional contributions to the scientific literature not directly related to this thesis, see Appendix \@ref(appendix-publications).

&nbsp;

### Presentations/Talks

<!--- Include links here for slides for each talk/poster --->

_"Identifying and triangulating all available evidence on the effect of blood lipids and statins on dementia outcomes"_ Accepted poster abstract, Alzheimer's Association International Conference 2021.

&nbsp;

_"medrxivr: A new tool for searching for and retrieving records and PDFs from the medRxiv preprint repository"_
Accepted oral presentation abstract, Cochrane Colloquium 2020 (note: event was cancelled due to the COVID-19 pandemic)

&nbsp;

_"On the shoulders of giants": advantages and challenges to building on established evidence synthesis packages, using the {robvis} package as a case study"_ Presentation to the  Evidence Synthesis and Meta-Analysis in R Conference (ESMARConf) 2021.

_"RoB 2.0: A revised tool to assess risk of bias in randomized trials"_ Webinar, co-presented with Dr. Theresa Moore as part of the Evidence Synthesis Ireland Methods Series.

### Software {#outputs-software}
**`medrxvir`**

An R package and associated `shiny` web application that allows users to easily search and retrieve bibliographic data from the medRxiv[@rawlinson2019] and bioRxiv[@sever2019] preprint repositories. See Chapter \@ref(sys-rev-tools-heading) for more details. Install a stable version of the package from the Comprehensive R Archive Network (CRAN), or alternatively install the development version from GitHub, using:

```{r, eval = FALSE}
# CRAN version
install.packages("medrxivr")

# Development version
devtools::install_github("mcguinlu/medrxivr")
```

&nbsp;

**`robvis`**

An R package and associated `shiny` web application that allows users to easily visualize the results of the risk-of-bias assessments performed as part of a systematic review. See Appendix \@ref(appendix-robvis) for more details. Install a stable version of the package from CRAN, or alternatively install the development version from GitHub, using:

``` {r, eval = FALSE}
# CRAN version
install.packages("robvis")

# Development version
devtools::install_github("mcguinlu/robvis")
```

&nbsp;

## Summary

This Chapter has provided background information on the core elements of the central research question, framed the research presented in this thesis in the context of an evidence synthesis framework, and described the contributions of the research in this thesis to the scientific literature.