### 2021-06-15 manuscripts/cprd/supp-materials.Rmd

---
title: 
date: "Supplementary Files updated: `r format(Sys.time(), '%d %B, %Y')`"
bibliography: references.bib
csl: vancouver.csl
pandoc_args: [       "-F", "./pandoc-word-pagebreak"     ]
output:
  bookdown::word_document2:
      toc: true
      number_sections: false
      toc_depth: 2
      reference_docx: "../../templates/word-styles-reference-01-no_line_num.docx"
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE) # Remove to knit with plots/data
options(scipen = 999)
source(here::here("R/doc_options.R"))
source(here::here("R/helper.R"))

#Load relevant libraries and datasets from the 'outputs' folder

knitr::read_chunk(here::here("R/04-Code-CPRD-Analysis.R"))
doc_type <- knitr::opts_knit$get('rmarkdown.pandoc.to')
```

\newpage

# Supplementary Materials {#supp-mat}

## Supplementary Material 1: Changes to protocol {#supp-mat-protocol}

Main changes include:

  * Use of a time-varying treatment indicator, to correctly classify time-at-risk.
  * Inclusion of additional covariates that are adjusted for in the main model.

\newpage

# Supplementary tables {#supp-tab}

## Supplementary Table 1: Definition of exposures and covariates

&nbsp;

<!----------------------------------------------------------------------------->
(ref:covariateDef-caption) Definition of covariates

```{r covariateDef-table, message=FALSE, results="asis", echo = FALSE}
```
<!----------------------------------------------------------------------------->

&nbsp;

\newpage

## Supplementary Table 2: Adherence and switching by drug class

&nbsp;

<!----------------------------------------------------------------------------->
(ref:cprdSSA-caption) cprdSSA

```{r cprdCharacteristics-table, message=FALSE, results="asis", include = FALSE}
```

```{r cprdSSA-table, message=FALSE, results="asis", echo = FALSE}
```
<!----------------------------------------------------------------------------->

&nbsp;

\newpage

## Supplementary Table 3: STROBE Reporting Guidelines {#supp-tab-strobe}

```{r child = 'supplementary_files/STROBE-Checklist.Rmd', eval = FALSE}
```

__To be completed__

\newpage

## Supplementary Table 4: Statin type prescribing by year {#supp-tab-sta-type}

&nbsp;

<!----------------------------------------------------------------------------->
(ref:statinTypeTable-caption) Statin prescription by year

```{r statinTypeTable-table, message=FALSE, results="asis", echo = FALSE}
```
<!----------------------------------------------------------------------------->

&nbsp;

\newpage

# Supplementary figures {#supp-fig}
## Supplementary Figure 1 {#supp-fig-outcome-decision}

```{r supp-fig-decision-tree,echo = FALSE, fig.cap="The algorithm used to choose between two diagnosis. This decision tree is adapted with permission from Walker _et al_ (2019).[@walker2019a]"}
knitr::include_graphics(here::here("figures/cprd-analysis/decision_tree.png"))
```

\newpage

## Supplementary Figure 2 {#supp-fig-attrition}
```{r supp-fig-attrition,echo = FALSE, fig.cap="Attrition of participants as the eligibility criteria were applied.", fig.width="60%"}
knitr::include_graphics(here::here("figures/cprd-analysis/cohort_attrition.png"))
```

\newpage

## Supplementary Figure 3 {#supp-fig-unadjusted}
```{r supp-fig-comparison,echo = FALSE, fig.cap="Association of any lipid regulating agent with a dementia or related outcome, adjusted only for age."}
knitr::include_graphics(here::here("figures/cprd-analysis/fp_unadjusted.jpeg"))
```

\newpage

## Supplementary Figure 4 {#supp-fig-entry}
```{r supp-fig-entry,echo = FALSE, fig.cap="Association of any lipid regulating agent with a dementia or related outcome, stratified by year of entry into the cohort. {Note: Venexia reccomend including some info on average length of follow-up for each group beside the HR - will be done before submission.}"}
knitr::include_graphics(here::here("figures/cprd-analysis/fp_cohort_entry.jpeg"))
```

\newpage

## Supplementary Figure 5 {#supp-fig-pregnancy}
```{r supp-fig-pregnancy,echo = FALSE, fig.cap="Association of any lipid regulating agent with a dementia or related outcome, removing women who were less than 55 years of age at index."}
knitr::include_graphics(here::here("figures/cprd-analysis/fp_pregnancy.jpeg"))
```

\newpage

## Supplementary Figure 6 {#supp-fig-sta-type}
```{r supp-fig-sta-type, echo = FALSE, fig.cap="Association of statins with a dementia or related outcome, stratified by statin lipophilic/hydrophilic properties."}
knitr::include_graphics(here::here("figures/cprd-analysis/fp_sta_type.jpeg"))
```

\newpage

## Supplementary Figure 7 {#supp-fig-backpain}
```{r supp-fig-backpain,echo = FALSE, fig.cap="Association of any lipid regulating agent with backpain and ischemic heart disease (IHD)."}
knitr::include_graphics(here::here("figures/cprd-analysis/fp_control_outcomes.jpeg"))
```

\newpage

# References for supplementary materials