### 2021-07-20 05-Chp-IPD-Analysis.Rmd
End of day

---
#########################################
# options for knitting a single chapter #
#########################################
output:
  bookdown::html_document2: default
  bookdown::word_document2: default
  bookdown::pdf_document2:
    template: templates/brief_template.tex
documentclass: book
---

```{block type='savequote', include=knitr::is_latex_output(), quote_author='(ref:intro-quote)', echo = FALSE}
Hold
```
(ref:intro-quote) --- Hold

# Individual participant data meta-analysis {#ipd-heading}
\minitoc <!-- this will include a mini table of contents-->

## Lay Summary

## Introduction

Is there a good way to visualise input of cohorts into an IPD analysis?

## Methods

As part of the IPD analysis, relevant cohort studies were identified from the sty and approached. In the first instance, the corresponding author on the main man

Talk about how different covariates were managed across cohorts,

Talk about missing data was handled <!-- IDEA See Sean's thesis for some additional detail on this -->

<!----------------------------------------------------------------------------->
&nbsp;

### Data sources

Several cohort studies were approach

<!-- Need exact number approach -->

As part of this Chapter, I will use individual patient data from a range of sources. These sources are described here in detail for reference.

Should also include a list of reasons why specific additional cohorts were not included - some like the EHR might be too big to get

Each should have:

* Description (including observation period, numbers, numbers with outcome, etc)
* Whether they are a known genetically at-risk cohort
*

<!----------------------------------------------------------------------------->
&nbsp;

#### Epic Norfolk

The European Prospective Investigation of Cancer - Norfolk is a  [@riboli1997; @riboli2002]

```{r tests, }

```

Different approaches to combining subgroups. [@fisher2017]

Several DPUK cohorts were excluded on the basis of a lack of exposure variables, including Cam-CAN (had cardiovascular category, but only contained blood pressure).

See

<!----------------------------------------------------------------------------->
&nbsp;

### Risk of bias assessment

Risk of bias assessment was performed for each of the included cohorts using the relevant tool described in Section \@ref(risk-of-bias).

<!--- Inlcude CPRD analysis here, but dou --->

<!----------------------------------------------------------------------------->
&nbsp;

### Data cleaning

<!----------------------------------------------------------------------------->
&nbsp;

## Discussion

Letters sent to all cohorts identified through the systematic review - will need quite a bit here on the numbers approached and how they responded if at all. Will also need to comment on the additional tactics used.

This is likely due to my junior position as a early career research combined with the impact

Range of reasons why data is not readily available. Privacy concerns, concerns around scoping or "parasitic" behaviour, and a lack of trust (i.e. primary researchers do not trust secondary researchers).<!-- TODO CITATION NEEDED: paper on trust as a big issue + paper on systematic review of barriers to data sharing -->

Unfortunately,

As part of this, an analysis of

### Reflections on data access

Describe your experience of trying to access the DPUK - while a great resources, frustrating at times. The portal would benefit from a centralised management system and

### Study within a review

In future, it could have been worth running a "study within a review", such as that used to examine whe

https://bmjopen.bmj.com/content/9/1/e025273