### 2021-09-13 tmp.Rmd

---
bibliography: bibliography/references.bib
csl: bibliography/nature.csl
output:
  bookdown::pdf_document2:
    template: templates/brief_template.tex
  bookdown::word_document2: 
      toc: false
      toc_depth: 3
      reference_docx: templates/word-styles-reference-01.docx
      number_sections: false 
  bookdown::html_document2: default
documentclass: book
---

```{r, echo = FALSE, warning=FALSE, message=FALSE}
source("R/doc_options.R")
source("R/helper.R")
knitr::read_chunk("R/03-Code-Systematic-Review.R")
doc_type <- knitr::opts_knit$get('rmarkdown.pandoc.to') # Info on knitting format
```

```{r primaryFigure1, include = FALSE}
```

<!-- Straight-forward random effects meta-analysis of binary exposures, using study defined measures of each. -->

In RCTs statin use was not associated

<!----------------------------------------------------------------------------->
(ref:statinsRCT-cap) __Random effects meta-analysis of statins on all-cause dementia:__

(ref:statinsRCT-scap) Random effects meta-analysis of statins on all-cause dementia

```{r statinsRCT, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:statinsRCT-cap)', out.width='100%', fig.scap='(ref:statinsRCT-scap)'}
knitr::include_graphics(file.path("figures/sys-rev/forest_statins_rct.png"))
```