### 2021-09-29 05-Chp-IPD-Analysis.Rmd
End of day:

---
output:
  bookdown::pdf_document2:
    template: templates/brief_template.tex
  bookdown::word_document2: default
  bookdown::html_document2: default
documentclass: book
---

```{block type='savequote', include=knitr::is_latex_output(), quote_author='(ref:intro-quote)', echo = FALSE}
Hold
```
(ref:intro-quote) --- Hold

# Individual participant data meta-analysis {#ipd-heading}
\minitoc <!-- this will include a mini table of contents-->

```{r, echo = FALSE, warning=FALSE, message=FALSE}
source("R/doc_options.R")
source("R/helper.R")
knitr::read_chunk("R/05-Code-IPD-Analysis.R")
doc_type <- knitr::opts_knit$get('rmarkdown.pandoc.to') # Info on knitting format
```

## Lay Summary

As part of a broader investigation into the relationship between lipid levels and dementia risk, I sought to

<!----------------------------------------------------------------------------->
&nbsp;

<!----------------------------------------------------------------------------->
&nbsp;

## Introduction

Is there a good way to visualise input of cohorts into an IPD analysis?

Justify analysis based on findings of systematic review - suggestion of variation of effect by age, specifically

<!----------------------------------------------------------------------------->
&nbsp;

## Methods

### Eligibility criteria

Eligible data sources were defined as those that contained lipid measurements and  ipids reported/available as a continuous measure, Data sources which were cross-sectional, either by design or due to the available data (e.g. a study conducted multiple waves, but onyl data from a single wave could be accessed) were excluded. Similarly, due to the limited time-frame, studies making use of population-level electronic health records, which often require an extensive project proposals in order to gain access to the data, were ineligible due to the time and cost involved in applying. The one exception to this was data from the CPRD, which I already had access to as part of the analysis reported in Chapter \ref(cprd-heading).

<!-- TODO Read this[@riley2020] -->

### Applying for data access

__Systematic review__

As part of the IPD analysis, eligible observational prospective cohort studies identified through the systematic review detailed in Chapters \@ref(sys-rev-heading) were approached for data access.

In the first instance, the first author on the main study report was emailed (see Appendix \@ref() for a copy of the text and documentation sent). If this did not elict a response, the last/corresponding author was contacted, on the basis that the first author may have been a more junior member.

examining the effect of lipid levels

__Dementia platform UK__

In addition, in an attempt to supplement the analysis with

In addition, to assess the results for  the Dementia Platform UK common-access procedure was utilised. In short, this approach is intended to make it easier to access data from existing dementia cohorts through a centralised application process.

&nbsp;<!----------------------------------------------------------------------->

### Risk of bias assessment

Risk of bias assessment was performed for each of the included cohorts using the risk-of-bias assessment tool for non-randomised studies of exposures introduced previously (for a more detailed discussion of this tool, see Section \@ref(rob-tools-nrse)).

<!-- Be clear on the different types of databases, e.g. those where users are followed up regularly vs those where participants are captured via EHR-->

&nbsp;<!----------------------------------------------------------------------->

### Missing data

<!-- TODO This was taken from Chapter 4, so may need to rephrase slightly -->
<!-- IDEA See Sean's thesis for some additional detail on this -->

In this analysis, missing data were handled using a multiple imputation approach. Variables with missing observations were identified, and 20 imputed datasets were created.[@sterne2009] Nominal variables with missing values were modelled using multinomial logistic regression, while continuous variables were modelled using linear regression. As per best practice, all variables used in the analytic model, including the outcome, were included in the imputation model.[@moons2006using] Imputation was performed using the MICE (Multiple Imputation by Chained Equations) using the XXX command in R.

Using multiple imputation to handle missing data is an alternative to a "complete case" approach,[@pigott2001review] where participants missing any covariate are dropped from the dataset. As a sensitivity analysis, I preformed and compared the results of both methods,[@hughes2019] to investigate the impact of multiple imputation on the results.

<!-- TODO Need to talk about how missing variables were handled across studies -->

<!----------------------------------------------------------------------------->
&nbsp;

### Analysis

_Individual patient data approached__

Two stage versus one-stage approach

A two-stage model was used out of necessity, with the one-stage model being precluded by the different datasets being in protected data silos.

__Statistical methods__

Under a two-stage approach, estimates for the effect of each lipid fraction on available dementia outcomes in each study were first calculated for each data source. A common model, adjusted for identical covariates, were obtained using a common model.

To aid interpretability of the outcome, all analyses were standardised by changes in 1-SD of the exposure variable. Clustering within studies was accounted for, given the evidence that ...[@abo-zaid2013]

Hazard ratios were used to quantify the effect of a 1-SD lower exposure to a lipid fraction on dementia outcomes. A discrete proportional hazards model was employed to account for the interval censoring introduced by design of the longitudinal cohort studies.[@wang2017]

The estimates from each data source were then combined using a random-effects meta-analysis

<!-- TODO talk about different methods - see here: https://onlinelibrary.wiley.com/doi/10.1002/sim.8516 -->
<!-- TODO Figure showing the outline of interval censored studies. -->

&nbsp;<!----------------------------------------------------------------------->

__Investigating effect of patient-level covariates__

In order to investigate the interaction of patient-level characteristics with lipid levels, interaction terms for lipid-covariate terms were included in the model above. These were extracted and synthesised using a random effects meta-analysis.

&nbsp;<!----------------------------------------------------------------------->

__Replicating published findings__

Where an included data source had previously been analysed and results for the association between lipid levels and dementia reported, these were compared to the results of this analysis. If discrepancies were identified, these were investigate.

__Combining published aggregate data and individual participant data__

Where a data source did not provided their individual participant data, but stratified by a patient characteristic of interest in the published data, the agregate date and individual data were combined.

<!-- TODO Need more on this method here -->

__Author statements on data availability__

As a sensitivity analysis, I examined the data sharing statements for each using a predefined categories used in a similar previous analysis.[@mcguinness2020DAScomparison]

<!-- TODO This should have a plot showing how many over time had good vs bad DAS -->

<!-- TODO In addition, need plot similar to study design over time in Chapter 3, showing cohort responses over time -->

&nbsp;<!----------------------------------------------------------------------->

## Results

### Data access

<!-- Need exact number from each approach, plus detailed exclusion reasons -->

Of the XXX studies to which I applied for data access, only three were eventually included in the analysis. Figure \@ref(fig:cohortFlowchart) details whether the cohorts eventually included in the review were identified by the systematic review or via DPUK portal.

In addition, the reasons for cohorts not being included in the analysis are presented. In summary, the requests for data from cohorts identified by the systematic review was characterised by a very low response rate. For the minority who did respond, common reasons given by corresponding authors for not sharing the data included that they no longer worked in the lab, had access to the data, or that they were currently or intended to perform a similar analysis as the one proposed.

For the streamlined DPUK process, where a dedicated project manager liaises with cohort owners on the applicants behalf, the overall response rate was higher, resulting in access to three cohorts. However, even using this dedicated access approach, a response was obtained for less than half of the approach cohorts

<!----------------------------------------------------------------------------->
```{r cohortFlowchartSetup, include = FALSE}
```

(ref:cohortFlowchart-cap) Flowchart of included cohorts, stratified by identification method (systematic review vs DPUK).

(ref:cohortFlowchart-scap) Flowchart

```{r cohortFlowchart, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:cohortFlowchart-cap)', out.width='100%', fig.scap='(ref:cohortFlowchart-scap)'}
knitr::include_graphics(file.path("figures/ipd/cohortFlowchart.png"))
```
<!----------------------------------------------------------------------------->

Few cohorts were included in both the DPUK and the systematic review sets of cohorts indicating that the DPUK.

As highlighted in the

&nbsp;<!----------------------------------------------------------------------->

### Included data sources

The four data sources used in this analysis are summarised in Table \@ref(tab:cohortOverview-table) and are described in detail in the following sections.

Of note, all data sources included in the analysis were based in the United Kingdom. This is likely due to the majority of included datasets coming from the Dementia Platform UK route \@ref(fig:cohortFlowchart), which as implied by the name, has a geographical focus on studies performed in Great Britain United Kingdom.

&nbsp;

<!----------------------------------------------------------------------------->
(ref:cohortOverview-caption) Summary of cohorts for which data were available

(ref:cohortOverview-scaption) Summary of included cohorts

```{r cohortOverview-table, message=FALSE, results="asis", echo = FALSE}
```
<!----------------------------------------------------------------------------->

&nbsp;

<!-- TODO Update table with SD for age, and types of outcomes, and follow-up, and check why number of events in EPIC is less than expected (5 vs 8)-->

<!-- TODO * Description (including observation period, numbers, numbers with outcome, etc) -->

&nbsp;<!----------------------------------------------------------------------->

#### Caerphilly Prospective Study

The Caerphilly <!-- TODO Check spelling --> Prospective Study is a longitudinal study of men in the

Cholesterol measures (total, LDL-c, HDL-c and triglyceride) were measured at baseline in 1979-1983. 
As the study population has aged, additional outcomes. Of particular relevance to this analysis, from Phase III (1989-1993) onwards, a battery of cognitive tests were introduced.

&nbsp;<!----------------------------------------------------------------------->

#### CPRD

The Clinical Practice Research Datalink (CPRD) is a large population-based, electronic health record (EHR) database.[@herrett2015] containing the primary care records for more than 10 million primary care patients in England, and is broadly representative of the UK population in terms of age, sex and ethnicity.[@herrett2015; @mathur2014]

The CPRD is introduced more fully in Chapter \@ref(). Briefly, a similar approach to the cohort definition as used in Chapter

Participants were included from the first date of lipid measurement, so no issues with immortal time bias as discussed in Chapter ...

Additionally, other number of participants is larger in this analysis as there is no restriction on the level which lipids should be in order to be included in the analysis.

the

&nbsp;<!----------------------------------------------------------------------->

#### Epic Norfolk

The European Prospective Investigation of Cancer - Norfolk is a  [@riboli1997; @riboli2002]

The added evidental value of the EPIC cohort is small, given the relatively small number of participants and the fact that the cohort has only 8 dementia events.

```{r tests}
```

&nbsp;<!----------------------------------------------------------------------->

#### Whitehall II

_The Whitehall II study is a prospective cohort study of 10 308 participants (70% men), aged 35–55 years and recruited between 1985 and 1989 from 20 London-based Civil service departments (https://www.ucl.ac.uk/whitehallII). Clinical examinations have been performed in 1991-1994, 1997-1999, 2002-2004, 2007-2009, 2012-2013, 2015-2016 with the data from circulating metabolomic traits and cognitive testing for the present study obtained from the 1997-1999 clinic phase._ - taken from EN ID: 2140

This data source was analysed in one of the included studies identified by the systematic review presented in Chapter \@ref(sys-rev-heading),[@tynkkynen2018] meaning that a comparison between the published result and the analysis reported here was was possible.

&nbsp;<!----------------------------------------------------------------------->

### Excluded data sources

#### Cohorts identified the systematic review but excluded from the IPD analysis

<!-- Should also include a list of reasons why specific additional cohorts were not included - some like the EHR might be too big to get -->

Electronic health record co

Lothian Birth Cohort - 1936

Taiwan, etc.

<!----------------------------------------------------------------------------->
&nbsp;

#### Cohorts identified in the DPUK but excluded from the IPD analysis

<!-- Should also include a list of reasons why specific additional cohorts were not included - some like the EHR might be too big to get -->

Several additional cohorts are present in the DPUK collection but were not approach with a request for data acccess. This was primarily due to the information on the online system indicating that the study was relatively new (and so was yet to collect >1 wave of data, as in the case of the XXXX cohort) or the online data dictionary indicating that the exposure variables of interest were not recorded.

Electronic health record co

LBC - 1936

Taiwan etc,

<!----------------------------------------------------------------------------->
&nbsp;

#### Cohorts providing data but ultimately excluded

As highlighted in Figure \@ref(fig:cohortFlowchart), several cohorts from the DPUK responding positively but on inspection of data provided these cohorts were excluded.

Several cohorts were excluded on the basis of a lack of exposure variables, including Cam-CAN (had cardiovascular category, but only contained blood pressure).

The reasons for exclusion of these cohorts for which data were provided is illustrated in Table \@ref(tab:dataExcluded-table).

__BRACE__

Likley had dementia at baseline. However, on further investigation of the cohort, it was discovered that

__Memento__

Excluded as criteria for entry was outpatients from memory clinic (unlikely to be dementia free at baseline)

__Generation Scotland__

Cross-sectional data only

__NICOLA__

NICOLA study - only had cross-sectional data

__ELSA__

Data provided but only had ever high cholesterol as a binary variable. Not compatible with

&nbsp;

<!----------------------------------------------------------------------------->
(ref:dataExcluded-caption) dataExcluded

(ref:dataExcluded-scaption) dataExcluded

```{r dataExcluded-table, message=FALSE, results="asis", echo = FALSE}
```
<!----------------------------------------------------------------------------->

&nbsp;

#### Cohorts approach but received no answer

A particularly frustrating in the case of cohorts which had a dedicated data access panel, for example, the Three City Study. Despite multiple attempts to contact the team, there was no response received.

These included studies with  a dedicated access process or data access committee. A cohort of particular interest, having been used in multiple studies identified by the systematic review is the Three City Study in France. This study was approach at multiple points over the curse of this analysis, and no answer was received.

This was also a problem for cohorts approach via the DPUK - half of the cohorts applied to did not respond to the application within a year.

<!-- TODO Will need some method to define how/from where the different cohorts were obtained. -->

<!----------------------------------------------------------------------------->
&nbsp;

### Data cleaning and harmonisation

Across all cohorts, data cleaning was performed in a similar manner, using commonly named variables, so that a single model could be applied using functional programming.

The one exception to this is the CPRD data, which was held in a different system to the rest.

The advantage of this approach is that it reduces the likelihood of errors in model mis-specification if needing to change variables names from cohort to cohort.

For all cohorts, the first lipid measurement was used for the exposure method,

For heterogeneity across the cohorts, the total time-at-risk was investigated.

<!-- TODO Create similar crude rate table as for CPRD analysis chapter -->

# Results

<!----------------------------------------------------------------------------->
(ref:tcDementiaMain-cap) __Random effects meta-analysis of estimates from included data sources:__

(ref:tcDementiaMain-scap) IPD analysis of total cholesterol and dementia

```{r tcDementiaMain, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:tcDementiaMain-cap)', out.width='100%', fig.scap='(ref:tcDementiaMain-scap)'}
knitr::include_graphics(file.path("figures/ipd/TC.png"))
```
<!----------------------------------------------------------------------------->

<!----------------------------------------------------------------------------->
&nbsp;

## Discussion

Useful citation for discussion[@levis2021]

### Limitations

#### Low response rate to request for data

<!-- TODO Plot showing when datasets requested were published and comparing this with the -->

This review had a low response rate to requests for data access.

While this is not unexpected, given that a review of IPD studies published between 1987 and 2015 found that fewer than half managed to obtained data from greater than 80% of studies, and that in many cases, the exact percentage of studies for which data were obtained was not accurately reported.
test
There are likely several reasons for this.

* Individual participant data meta-analysis including studies other than randomised controlled trials have less success in obtaining individual participant data from studies.[@nevitt2017a]

* While there is no

For many other reasons also - if

A potential further reason was highlighted during my attendance at the

Letters sent to all cohorts identified through the systematic review 
<!-- TODO will need quite a bit here on the numbers approached and how they responded if at all. Will also need to comment on the additional tactics used.  -->

This is likely due to my junior position as a early career research combined with the impact

Range of reasons why data is not readily available. Privacy concerns, concerns around scoping or "parasitic" behaviour, and a lack of trust (i.e. primary researchers do not trust secondary researchers).<!-- TODO CITATION NEEDED: paper on trust as a big issue + paper on systematic review of barriers to data sharing -->

Unfortunately,

The low response rate and the

Whether or not to press ahead with an IPD analysis in the absence of all (or even most) data is a personal decision, and some have highlighted where they decided not to pursue an IPD study.[@jaspers2014]

<!-- However, fewer than half of systematic reviews with IPDMA, published between 1987 and 2015, retrieved data from at least 80% of relevant studies and from at least 80% of relevant participants [16]. The number of IPDMAs increased over this period [17], but data retrieval rates remained unchanged [16, 18]. Inability to include eligible studies compromises the systematic review’s purpose, decreases study power, and leads to healthcare decisions based on an incomplete, potentially biased data sample (studies with available data may differ from those whose data are not available) [10, 19, 20]. However, analysis combining individual and study level data may mitigate these effects [3, 21]. from [@ventresca2020a] -->

For the purposes of this thesis, conducting the IPD was useful as it afforded me the opportunity to experience the methodology.

On a personal level, given the

In addition, it allowed me to analysed two addition previously unanalysed cohort studies which were then utilised in the triangulation study detailed in Chapter \@ref(tri-heading).

#### Risk of bias assessment

There is some concerns about performing risk of bias assessments on your own analysis, and so two secondary reviewers

#### Opportunistic inclusion of CPRD data

&nbsp;<!----------------------------------------------------------------------->

### Strengths

While this analysis did not manage to systematically obtain and analyse a large proporiton of identified data, it did enable an analysis of two previously unanalysed datasets - the CaPS study and the - providing additional data points in

it also attempted to make use

&nbsp;<!----------------------------------------------------------------------->

### Data access

Couple of sections

* Problems with getting data from studies
* Problems with current methods of tracking people down via . Particularly when studies were older, several cases of the address/contact details for the corresponding author being out of date. Matches with decrease in availability as the studies get older.

__Many authors claim that their data can be made “available on request”, despite previous work establishing that these statements are demonstrably untrue in the majority of cases—that when data is requested, it is not actually made available. [@krawczyk2012; @miyakawa2020; @naudet2018] Additionally, previous work found that the availability of data “available on request” declines with article age, indicating that this approach is not a valid long term option for data sharing.[@vines2014]__ <!-- TODO taken from DAS paper, so will need to change language -->

Qualitatively, investigation of a subset of cohorts on an _ad-hoc_ basis reveal that several corresponding/first authors were no longer at the same institution as when the study was reported, and as a result, were unlikely to have access to the institutional email address listed on the study publication. All of these are arguments towards better reporting of data availability, ownership and

Even with the help of the streamlined application process afforded by the DPUK, accessing sufficient data was a challenge. Sharing data has several legal, ethical and logistical challenges, and can often taken several years to obtain. This is in theory what the DPUK was built to address. However, despite claims to a streamlined process, the response rate among cohorts a year after application was 50%. In addition, some cohorts responded that that the proposed study question was already under investigation by another group, and that they would not share the data on this basis. In light of this, a centralised database of ongoing analysis being performed in DPUK would be of enormous help. Finally, the DPUK should more clearly distinguish between those cohorts that are "DPUK native" versus externally hosted, as the response time for externally hosted cohorts is likely to be much longer.

&nbsp;<!----------------------------------------------------------------------->

### Study within a review

In future, it could have been worth running a "study within a review", such as that used to examine the best method [@godolphin2019a] to identify and access study data.

Options could include

Previous SWARs have been registered as protocols, but have not yet reported!

&nbsp;<!----------------------------------------------------------------------->

### Reflections on the process

#### Aims/Resources

May have been overly ambitious

While it was disappointing that such a small proportion of the available data were obtained for analysis, in terms of project completion, it is probably a good thing.

Given the resource intensity of data cleaning and harmonisation,

&nbsp;<!----------------------------------------------------------------------->

\newpage

## References