### 2021-09-29 front-and-back-matter/_00-introduction.Rmd

# Covering material {-}

\adjustmtc 
<!-- For PDF output, we must include this LaTeX command after unnumbered headings, otherwise the numbers in the mini table of contents will be incorrect -->

```{r word-count, echo=FALSE, warning=FALSE, message=FALSE}
options(scipen=999)

source(here::here("R/helper.R"))

words <- get_words()

table_words <- table_words()

write.csv(file = here::here("data","words","words.csv"),
          x = words,
          col.names = FALSE,
          row.names = FALSE)

start_date <- as.Date("20201124", format="%Y%m%d")

max_date <- as.Date(max(words$date), format="%Y%m%d")

words_today <- words$words[which(as.Date(words$date, format="%Y%m%d")==max_date)]

date_diff <- as.numeric((max_date) - start_date)

target <- (date_diff*97) + 5580

words_behind <- words$words[which(words$date == max(words$date))] - target

words_new_today <- words$words[which(words$date == max(words$date))] - words$words[length(words$words)-1]

target_df <- data.frame(date = c(start_date,as.Date("20211116", format="%Y%m%d")),
                           words = c(5580, 40000))

words$date <- as.Date(words$date, format="%Y%m%d")

```

Word count: `r words_today`

Table words: `r table_words`

Days: `r date_diff`

Words behind: `r format(words_behind, scientific = F)`

Words today: `r words_new_today`

&nbsp;

&nbsp;

```{r word-count-fig, echo=FALSE, warning=FALSE, message=FALSE}
ggplot(mapping = aes(x=date,y=words)) +
  geom_line(data = target_df, colour = "black", linetype = 2) +
  geom_hline(yintercept = 40000, colour = "black", linetype = 2) +
  geom_vline(xintercept = as.numeric(as.Date("20211116", format="%Y%m%d")), colour = "black", linetype = 2) +
  geom_vline(xintercept = as.numeric(as.Date("20211216", format="%Y%m%d")), colour = "red") +
  scale_x_date(limits = c(as.Date("20201124", format="%Y%m%d"),as.Date("20211231", format="%Y%m%d"))) +
  geom_step(data = words, colour = "red") +
  geom_text(aes(x = as.Date("20210910", format="%Y%m%d"), y = 30000, label = "Tables"))

```