### 2021-10-13 manuscripts/cprd/supp-materials.Rmd

---
title: 
date: "Supplementary Files updated: `r format(Sys.time(), '%d %B, %Y')`"
bibliography: references.bib
csl: vancouver.csl
pandoc_args: [       "-F", "./pandoc-word-pagebreak"     ]
output:
  bookdown::word_document2:
      toc: true
      toc_depth: 2
      number_sections: false
      reference_docx: "../../templates/word-styles-reference-01.docx"
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE) # Remove to knit with plots/data
source(here::here("R/helper.R"))
#Load relevant libraries and datasets from the 'outputs' folder
knitr::read_chunk(here::here("R/05-Code-CPRD-Analysis.R"))
doc_type <- knitr::opts_knit$get('rmarkdown.pandoc.to')
```

\newpage

# Supplementary Materials {#supp-mat}

## Supplementary Material 1: Changes to protocol {#supp-mat-protocol}

Any changes made to the approved protocol are "Minor amendments" as per the ISAC criteria, specifically falling under the category of "Additional methods to further control for confounding or sensitivity analysis provided these are to be reported as secondary to the main findings".

Main changes include:

-   Use of a time-varying treatment indicator, to correctly classify time-at-risk.
-   Inclusion of additional covariates that are adjusted for in the main model.

\newpage

## Supplementary Material 2: RECORD Reporting Guidelines {#supp-tab-strobe}

```{r child = 'supplementary_files/STROBE-Checklist.Rmd', eval = FALSE}
```

A copy of the RECORD Checklist[@nicholls2016record] is provided below.

**To be copied in**

\newpage

# Supplementary tables {#supp-tab}

## Supplementary Table 1: Definition of exposures and covariates

(ref:covariateDef-caption) Definition of covariates adjusted for in the Cox PR model.

The code lists used to define covariates adjusted for in the fully-adjusted model (see Supplementary Table \@ref(tab:covariateDef-table)) were originally created for use in a previous analysis.[@walker2020] Some code lists were built on or adapted from previous published work,[@khan2010; @taylor2016; @wright2017] and these are noted in the table.

```{r covariateDef-table, echo = FALSE}
```

\newpage

## Supplementary Table 2: Adherence and switching by drug class

(ref:cprdSSA-caption) Adherence and switching by drug class.

<!----------------------------------------------------------------------------->

```{r cprdCharacteristics-table, message=FALSE, results="asis", include = FALSE}
```

```{r cprdSSA-table, message=FALSE, echo = FALSE}
```

<!----------------------------------------------------------------------------->

\newpage

# Supplementary figures {#supp-fig}

## Supplementary Figure 1 {#supp-fig-outcome-decision}

```{r supp-fig-decision-tree,echo = FALSE, fig.cap="The algorithm used to choose between two diagnosis. This decision tree is adapted with permission from Walker _et al_ (2020).[@walker2020]"}
knitr::include_graphics(here::here("figures/cprd-analysis/decision_tree.png"))
```

\newpage

## Supplementary Figure 2 {#supp-fig-attrition}

```{r supp-fig-attrition,echo = FALSE, fig.cap="Attrition of participants as the eligibility criteria were applied.", out.width='60%'}
knitr::include_graphics(here::here("figures/cprd-analysis/cohort_attrition.png"))
```

\newpage

```{r unadjustedComparison, include = FALSE}
```

```{r pregnancy, include = FALSE}
```

```{r statinType, include = FALSE}
```

```{r controlOutcomesText, include = FALSE}
```

```{r controlOutcomes, include = FALSE}
```

## Supplementary Figure 3 {#supp-fig-unadjusted}

```{r supp-fig-comparison,echo = FALSE, fig.cap="Association of any lipid regulating agent with a dementia or related outcome using three models adjusted for age, age and sex, and all covariates respectively."}
knitr::include_graphics(here::here("figures/cprd-analysis/forester_unadjusted.png"))
```

\newpage

## Supplementary Figure 4 {#supp-fig-pregnancy}

```{r supp-fig-pregnancy,echo = FALSE, fig.cap="Association of any lipid regulating agent with a dementia or related outcome, removing participants who were less than 55 years of age at index."}
knitr::include_graphics(here::here("figures/cprd-analysis/forester_pregnancy.png"))
```

\newpage

## Supplementary Figure 5 {#supp-fig-sta-type}

```{r supp-fig-sta-type, echo = FALSE, fig.cap="Association of statins with a dementia or related outcome, stratified by statin lipophilic/hydrophilic properties."}
knitr::include_graphics(here::here("figures/cprd-analysis/forester_sta_type.png"))
```

\newpage

## Supplementary Figure 6 {#supp-fig-backpain}

```{r supp-fig-backpain,echo = FALSE, fig.cap="Association of any lipid regulating agent with backpain, ischemic heart disease (IHD), and Type 2 diabetes."}
knitr::include_graphics(here::here("figures/cprd-analysis/forester_control_outcomes.png"))
```

\newpage

# References for supplementary materials