### 2021-11-19 Disc2.Rmd

## Summary of principal findings

No impact of lipid factions on dementia outcomes, except for triglycerides on vascular dementia outcomes

Non-randomised studies of interventions demonstrated a slight protective effect of statins. MR studies and RCT did not replicate the findings. However, the risk of bias in many of these studies was high, and when included in a bias adjusted framework.

Statins are by far the most studied LRA in relation to dementia outcomes, while there was an absence of information on the

Including preprints adds valuable additional evidence to a synthesis.

Reporting bias is particularly prevalent in NRSE - many studies explicitly state that they choose not to report

Across the studies, seems to be some evidence of triglycerides of VaD