### 2021-12-09 front-and-back-matter/abbreviations.tex

% do not edit by hand - add to abbreviations.csv instead
% First parameter can be changed eg to "Glossary" or something.
% Second parameter is the max length of bold terms.
\begin{mclistof}{List of Abbreviations}{3.2cm}
\item[AC] Agreement coefficient
\item[ACE] Acetylcholinesterase
\item[AD] Alzheimer's disease
\item[ADGC] Alzheimer’s Disease Genetics Consortium
\item[API] Application programming interface
\item[ApoE] Apolipoprotein E
\item[BAS] Bile acid sequestrants
\item[BMI] Body mass index
\item[CAD] Coronary arterial disease
\item[CaPS]  Caerphilly Prospective Study
\item[CBS] Coronary bypass surgery
\item[CCI] Charlson co-morbidity index
\item[CIND] Cognitive impairment not dementia
\item[CKD] Chronic kidney disease
\item[CPRD] Clinical Practice Research Datalink
\item[CRAN] Comprehensive R Archive Network
\item[CSV] Comma separated value
\item[CVD] Cardiovascular disease
\item[DOI] Digitial object identifier
\item[DPUK] Dementia Platform UK
\item[DSM] Diagnostic and Statistical Manual of Mental Disorders
\item[EHR] Electronic health record
\item[EPIC] European Prospective Investigation of Cancer 
\item[FAGs] Fatty acid groups
\item[GLGC] Global Lipids Genetic Consortium
\item[GWAS]  Genome wide association studies
\item[HDL] High density dipoprotein
\item[HES] Hospital episode statistics
\item[HMG-CoA] 3-hydroxy-3-methylglutaryl-coenzyme-A
\item[HMGCR] HMG-CoA reductase
\item[HR] Hazard ratio
\item[ICD] International Classification of Diseases
\item[IGAP] International Genomics of Alzheimer's Project
\item[IHD]  Ischemic heart disease
\item[IMD] Index of multiple deprivation
\item[IPD] Individual participant data
\item[IV] Instrumental variable
\item[JUPITER] Justification for the Use of Statin in Prevention: An Intervention Trial Evaluating Rosuvastatin
\item[LDL-c] Low density lipoprotein cholesterol
\item[LRA] Lipid regulating agent
\item[MCI] Mild cognitive impairment
\item[MICE] Multiple imputation by chained equations
\item[MMSE] Mini Mental State Exam
\item[MoCA] Montreal Cognitive Assessement
\item[MR] Mendelian randomization
\item[NAG] Nicotinic acid groups
\item[NINDS-ADRA] National Institute of Neurological Disorders and Stroke - Alzheimer's Disease and Related Disorders Association
\item[NINDS-AIREN] National Institute of Neurological Disorders and Stroke - Association Internationale pour la Recherche en l'Enseignement en Neurosciences
\item[NOS] Newcastle-Ottowa Scale
\item[NRSE] Non-randomised studies of exposures
\item[NRSI] Non-randomised studies of interventions
\item[OR] Odds ratio
\item[PAD] Peripheral arterial disease
\item[PCSK9i] PCSK9 inhibitors
\item[PDF] Portable document format
\item[PR] Proportional hazards
\item[PROSPER] Prospective Study of Pravastatin in the Elderly
\item[PYAR] Participant-years-at-risk
\item[RCT] Randomised controlled trial
\item[REML] Restricted maximum likelihood 
\item[RoB-ME] Risk of bias due to missing evidence
\item[RR] Risk/rate ratio
\item[SD] Standard deviation
\item[SNP] Single nucleotide polymorphism
\item[TC] Total cholesterol
\item[TG] Triglycerides
\item[VaD] Vascular dementia
\end{mclistof}