### 2021-12-14 front-and-back-matter/_acknowledgements.Rmd
Update acknowledgements and dedication

<!-- TODO UPDATE and check all name spellings -->

__Professional__

There are many people without whose input and support this work would have been severely diminised:

Thank Julian separately -

First and foremost, I would like to thank my supervisors - Julian Higgins, Yoav Ben-Shlomo, Liz Coulthard and George Davey-Smith - for all their support over the past three years. Thanks also to Rupert Payne and Emma Anderson for their engaging discussion and useful feedback during my annual progress reviews. Sharen Hockey-O'Keffe helped me at numerous points throughout my PhD, and I was very lucky to receive her support. Similarly, Mark Newbury of the Dementia Platform UK worked tirelessly to support my applications for data access - I am extremely grateful for his help and commitment in chasing cohort owners.

Thanks must also go to those who contributed their time and energy to my research projects: Matthew Lee, Lena Schmidt, Alexandra McAleenan, and Vincent Cheng. Francesca Spiga kindly shared an advance copy of her review of tools to assess Mendelian randomisation studies, and Rebecca Turner shared her code for bias-/indirectness-adjusted meta-analyses that underpins the quantitative triangulation framework presented in Chapter \@ref(tri-heading).

A particular thank you to Emily Hennessy, with whom I co-wrote the guidance paper for preparing primary research for incorporation into systematic reviews. Similarly, thanks to Athena Sheppard for her support in assessing the impact of journal polices on open-data sharing.

Thanks also to all those who raised issues or otherwise provided feedback on the open source tools built as part of this thesis:
<<github>> A special mention must go to Randall Boyes and Alex J Fowler for their input on the `robvis` tool.

I have been supported throughout my studies by an NIHR Doctoral Research Fellowship (DRF-2018-11-ST2-048). The views expressed in this thesis are my own and do not necessarily represent those of the NIHR or the Department of Health and Social Care.

\bigskip

__Personal__

This thesis would not have been possible with Ciara Gardiner. Lots of people say this about someone at this point in the acknowledgements, and I'm always sceptical of whether it is true, so to avoid any doubt I mean it here quite literally.

Ciara, I'd need (at least) another <<words>> words to accurately describe all the ways in which you supported me through this endeavour. It can't have been fun being locked-down in a small flat with a grumpy and stressed PhD student, nor trying to support the same student remotely from Dublin. I talk about lessons learnt during this thesis in the Discussion, but the biggest takeaway for me is that I am incredibly lucky to have found you and to get to take on life while holding your hand. Thank you and I love you.

Thanks to Mum and Dad for your never-ending love and support, and everything they have done to get me to point where doing and submitting a thesis.

Thanks to Vivienne and Dad for giving up a substantial amount of time to proof-reading.

On a personal note, there are several people without whom I would not have survived the process of writing a dissertation during COVID times. Ben, Doug, Christin - thank you!

\bigskip

\begin{flushright}
Canynge Hall, Bristol \\\
1 December 2021
\end{flushright}