### 2021-12-16 front-and-back-matter/_acknowledgements.Rmd

<!-- TODO UPDATE and check all name spellings -->

__Professional__

First and foremost, I must thank my primary supervisor, Julian Higgins, for his tireless support and encouragement. I'd also like to thank my other supervisors - Yoav Ben-Shlomo, Liz Coulthard and George Davey-Smith - for all their input over the past three years. Thanks also to Rupert Payne and Emma Anderson for their engaging discussion and useful feedback during my annual progress reviews.

I am extremely grateful to all those who contributed their time and energy to my research projects: Athena Sheppard, Emily Hennessy, Matthew Lee, Lena Schmidt, Alexandra McAleenan, Vincent Cheng, Francesca Spiga, Sarah Dawson and Rebecca Turner. A special mention goes to Mark Newbury of the Dementia Platform UK, who worked tirelessly to support my applications for data access - I am extremely grateful for his help and commitment in chasing cohort owners.

I also want to thank the many people who raised issues or otherwise provided feedback on the open source tools built as part of this thesis - it's been a joy to watch others build on my work. A special mention must go to Randall Boyes and Alex J Fowler for their input/work on the `robvis` tool.

\bigskip

__Personal__

Ciara, I'd need (at least) another <<words>> words to accurately describe all the ways in which you supported me through this endeavour. It can't have been fun being locked-down in a small flat with a grumpy and stressed PhD student, nor trying to support the same student remotely from Dublin. In the Discussion, I talk about lessons learned during this thesis, but the biggest takeaway of this whole experience for me is that I am incredibly lucky to have found you and to get to take on life while holding your hand. Thank you and I love you.

Mum and Dad, thank you for your never-ending love and support. Everything I am comes from you, and you've done so much to get me to the point where a PhD thesis seemed possible.

There are several people without whom I would not have survived the process of writing a dissertation during COVID times. Ben, Doug, Christin - thank you!

Finally, a big thanks to Vivienne Gardiner, who sacrificed a substantial amount of time to proof-reading this thesis.

\bigskip

__Funding__

I have been supported throughout my studies by an NIHR Doctoral Research Fellowship (DRF-2018-11-ST2-048). The views expressed in this thesis are my own and do not necessarily represent those of the NIHR or the Department of Health and Social Care.

\bigskip

\begin{flushright}
Canynge Hall, Bristol \\\
1 December 2021
\end{flushright}