### 2021-12-16 front-and-back-matter/99-appendices.Rmd
End of day:

---
output:
  word_document: default
  html_document: default
---
`r if(knitr:::is_latex_output()) '\\startappendices'`

`r if(!knitr:::is_latex_output()) '# (APPENDIX) Appendix {-}'`

```{r, echo = FALSE, warning=FALSE, message=FALSE}
source(here::here("R/helper.R"))
knitr::read_chunk(here::here("R/appendix.R"))
doc_type <- knitr::opts_knit$get('rmarkdown.pandoc.to') # Info on knitting format
```

# Appendices by Chapter {#chapter-appendix-heading}

\minitoc <!-- this will include a mini table of contents-->

\newpage

## Chapter 1 {#appendix-into}

### Publications beyond the scope of this thesis {#appendix-publications}

In addition to publications arising directly from this thesis (described in Section \@ref(literature-contributions) and included in Appendix \@ref(published-papers)), I have been involved in a number of external research projects. Taken together, these additional projects provided extensive experience in evidence synthesis methods and in collaborating across large teams.

During 2020, I was involved in the updating of the Preferred Reporting Items for Systematic reviews and Meta-Analyses (PRISMA) statement. My contributions were focused on the data sharing aspects of the guidelines and on building the PRISMA2020 interactive web application. Publications arising from this work are listed below:

*	*Page, M.J., Moher, D., Bossuyt, P., Boutron, I., Hoffmann, T., Mulrow, C., Shamseer, L., Tetzlaff, J., Akl, E., Brennan, S.E., Chou, R., Glanville, J., Grimshaw, J., Hróbjartsson, A., Lalu, M.M., Li, T., Loder, E., Mayo-Wilson, E., McDonald, S., __McGuinness, L.A.__, Stewart, L., Thomas, J., Tricco, A., Welch, V.A., Whiting, P., McKenzie, J. (2021). PRISMA 2020 statement: updated guidelines for reporting systematic reviews and meta-analyses. BMJ 372:n71. DOI:  [10.1136/bmj.n71](https://doi.org/10.1136/bmj.n71)*

*	*Page, M.J., Moher, D., Bossuyt, P., Boutron, I., Hoffmann, T., Mulrow, C., Shamseer, L., Tetzlaff, J., Akl, E., Brennan, S.E., Chou, R., Glanville, J., Grimshaw, J., Hróbjartsson, A., Lalu, M.M., Li, T., Loder, E., Mayo-Wilson, E., McDonald, S., __McGuinness, L.A.__, Stewart, L., Thomas, J., Tricco, A., Welch, V.A., Whiting, P., McKenzie, J. (2021). PRISMA 2020 explanation and elaboration: updated guidance and exemplars for reporting systematic reviews. BMJ 372:n160. DOI: [10.1136/bmj.n160](https://doi.org/10.1136/bmj.n160)*

* *Haddaway, N. R., Page, M. J., Pritchard, C. C., & McGuinness, L. A. (2021). PRISMA2020: An R package and Shiny app for producing PRISMA 2020-compliant flow diagrams, with interactivity for optimised digital transparency and Open Synthesis. MedRxiv, DOI: [10.1101/2021.07.14.21260492](https://doi.org/10.1101/2021.07.14.21260492)*

&nbsp;

In response to the COVID-19 pandemic, I worked with a team of evidence synthesists to develop a literature searching/screening pipeline. This work underpinned a living systematic review on impact of the COVID-19 pandemic on self-harm and suicidal behaviour. Subsequently, the pipeline enabled rapid reviews of related topics, such as an analysis of the available evidence by population.

*	*John, A., Okolie, C., Eyles, E., Webb, R.T., Schmidt, L., __McGuinness, L.A.__, Olorisade, B.K., Arensman, E., Hawton, K., Kapur, N. and Moran, P., 2020. The impact of the COVID-19 pandemic on self-harm and suicidal behaviour: a living systematic review. F1000Research, 9(1097). DOI: &nbsp; [10.12688/f1000research.25522.2](https://doi.org/10.12688/f1000research.25522.2)*

*	*John, A., Eyles, E., __McGuinness, L.A.__, Okolie, C., Olorisade, B.K., Schmidt, L., Webb, R.T., Arensman, E., Hawton, K., Kapur, N. and Moran, P., 2020. The impact of the COVID-19 pandemic on self-harm and suicidal behaviour: protocol for a living systematic review. F1000Research, 9(644), p.644. DOI: [10.12688/f1000research.24274.1](https://doi.org/10.12688/f1000research.24274.1)*

* *Eyles, E., Moran, P., Okolie, C., Dekel, D., Macleod-Hall, C., Webb, R. T., Schmidt, L., Knipe, D., Sinyor, M., __McGuinness, L. A.__, Arensman, E., Hawton, K., O’Connor, R. C., Kapur, N., O’Neill, S., Olorisade, B., Cheng, H.-Y., Higgins, J. P. T., John, A., & Gunnell, D. (2021). Systematic review of the impact of the COVID-19 pandemic on suicidal behaviour amongst health and social care workers across the world. Journal of Affective Disorders Reports, 6, 100271. DOI: [10.1016/j.jadr.2021.100271](https://doi.org/10.1016/j.jadr.2021.100271)*

&nbsp;

Finally, I have acted as part of a large review team for a number of other evidence synthesis exercises:

*	*Grant, M. C., Geoghegan, L., Arbyn, M., Mohammed, Z., __McGuinness, L.A.__, Clarke, E. L., & Wade, R. (2020). The Prevalence of Symptoms in 24,410 Adults Infected by the Novel Coronavirus (SARS-CoV-2; COVID-19): A Systematic Review and Meta-Analysis of 148 Studies from 9 Countries. PLOS ONE. DOI: [10.1371/journal.pone.0234765](https://doi.org/10.1371/journal.pone.0234765)*

* *Cheng, H.Y., __McGuinness, L.A.__, Elbers, R.G., MacArthur, G.J., Taylor, A., McAleenan, A., Dawson, S., López-López, J.A., Higgins, J.P., Cowlishaw, S. and Lingford-Hughes, A. (2020). Treatment interventions to maintain abstinence from alcohol in primary care: systematic review and network meta-analysis. BMJ, 371. DOI: [10.1136/bmj.m3934](https://doi.org/10.1136/bmj.m3934)*

*	*Schmidt, L., Olorisade, B.K., __McGuinness, L.A.__, Thomas, J. and Higgins, J.P. (2020). Data extraction methods for systematic review (semi) automation. F1000Research, 10:401. DOI: [10.12688/f1000research.51117.1](https://doi.org/10.12688/f1000research.51117.1)*

&nbsp;<!----------------------------------------------------------------------->

## Chapter 2 {#appendix-sys-rev-tools}

### Code for publication rate analysis {#appendix-tools-pub}

medRxiv automatically links preprints to their published counterparts and provides the DOI of the published article as part of the metadata on a given preprint. Using `medrxivr` to access this data, I could assess the two-year publication rate of preprints originally posted to medRxiv in July 2019. The code used for this analysis is presented below:

&nbsp;

```{r, eval=FALSE}
# Load relevant packages
library(medrxivr)
library(dplyr)

# Use snapshot of published status from July 2021
  mx_search(
    mx_snapshot("ccedfb8a44304b9fba4e3ba518a8ce4ed2294770"),
    query = "*",
    # Limit to those 129 records published in July 2019
    from_date = "2019-07-01",
    to_date = "2019-08-01"
  ) %>%

# Create indicator to show which records have been published
  mutate(Status = ifelse(published == "NA",
                          "Not published",
                          "Published")) %>%

# Group by indicator variable ("Published"/"Not published")
  group_by(Status) %>%

# Get total number per group, and percentage of total (N=129)
  summarise(N = n(),
            Percent = comma(n() / 129 * 100)) 
```

&nbsp;

As shown in Table \@ref(tab:medRes-table), two-thirds of preprints were published within two years of their submission to medRxiv.

&nbsp;

<!----------------------------------------------------------------------------->
(ref:medRes-caption) __Publication status of medRxiv preprints__ - The publication status of medRxiv preprints deposited in July 2019 was assessed in July 2021 (allowing for a two-year lag) and found that a substantial proportion remained unpublished.

(ref:medRes-scaption) Publication status of medRxiv preprints

```{r medRes-table, message=FALSE, results="asis", echo = FALSE}
```
<!----------------------------------------------------------------------------->

&nbsp;

## Chapter 3 {#appendix-sys-rev}

### Deviations from protocol {#appendix-sys-rev-deviations}

There were a number of deviations from the pre-registered protocol for the systematic review reported in Chapters \@ref(sys-rev-methods-heading) & \@ref(sys-rev-results-heading):[@mcguinnessluke2020]

* I had originally intended to examine both dementia and mild cognitive impairment (MCI) as outcomes of interest. However, given the large number of relevant records for the dementia outcomes, I did not include MCI as an outcome.

* For Mendelian randomisation analyses, I had originally intended to use a panel of experts to assess the risk of bias in this study design. However, through private correspondence with the authors of a review of risk-of-bias assessment tools for Mendelian randomisation studies (since published[@spiga2021]), I became aware of the Mamluk _et al_ tool. Given this tool was considered to be the best available for assessing Mendelian randomisation studies and that it followed a domain-based approach, enabling comparison with tools for other study designs, I decided to use it in the review.

* I performed an additional assessment of the risk of bias due to missing evidence using the ROB-ME tool. This analysis was not planned because the tool did not exist when the protocol was written.

* I had originally intended to assess sources of funding alongside bias. However, on further reading,[@sterne2013] I learned that any bias introduced by sources of funding would operate through one of the existing domains of the risk-of-bias tools. As such, I did not consider funding sources separately.

 * Finally, I performed a dose response meta-analysis to assess the evidence for a non-linear relationship between lipids and dementia outcomes. This marked a departure from the published protocol, as I had not expected the number of studies reporting dementia risk across lipid sub-categories to be sufficient for a dose response analysis.

&nbsp;<!----------------------------------------------------------------------->

### Web of Science databases searched {#appendix-wos-databases}

(ref:wosDatabases-caption) __Summary of Web of Science databases searched__ - The contents of the Web of Science core collection can vary depending on an institution's subscription, and so detailing the exact databases and ranges searched is important.

(ref:wosDatabases-scaption) Summary of Web of Science databases searched

```{r wosDatabases-table, message=FALSE, results="asis", echo = FALSE}
```

&nbsp;<!----------------------------------------------------------------------->

### Search strategy {#appendix-search-strategy}

An overview of the number of hits per database, in addition to the full line-by-line search results for each database, are presented below.

(ref:searchOverviewAppendix-caption) __Summary of number of records identified in each database searched__ - The full line-by-line search results for each database, are presented in subsequent tables. Note there was substantial duplication across databases, resulting in several thousand duplicates being removed prior to title and abstract screening.

(ref:searchOverviewAppendix-scaption) Summary of number of records identified in each database searched

```{r searchOverviewAppendix-table, message=FALSE, results="asis", echo = FALSE}
```

&nbsp;<!----------------------------------------------------------------------->

(ref:searchHitsMedline-caption) __MEDLINE search strategy__ - Full list of search terms used to identify relevant records in the MEDLINE bibliographic database. The number of records returned per search term is also displayed.

(ref:searchHitsMedline-scaption) MEDLINE search strategy

```{r searchHitsMedline-table, message=FALSE, results="asis", echo = FALSE}
```

&nbsp;

(ref:searchHitsEmbase-caption) __Embase search strategy__ - Full list of search terms used to identify relevant records in the Embase bibliographic database. The number of records returned per search term is also displayed.

(ref:searchHitsEmbase-scaption) Embase search strategy

```{r searchHitsEmbase-table, message=FALSE, results="asis", echo = FALSE}
```

&nbsp;

(ref:searchHitsPI-caption) __PyscINFO search strategy__ - Full list of search terms used to identify relevant records in the PyscINFO bibliographic database. The number of records returned per search term is also displayed.

(ref:searchHitsPI-scaption) PyscINFO search strategy

```{r searchHitsPI-table, message=FALSE, results="asis", echo = FALSE}
```

&nbsp;

(ref:searchHitsCentral-caption) __CENTRAL search strategy__ - Full list of search terms used to identify relevant records in the CENTRAL bibliographic database. The number of records returned per search term is also displayed. Note that at the last step, I applied a manual filter via the CENTRAL web interface to limit the search to primary studies.

(ref:searchHitsCentral-scaption) CENTRAL search strategy

```{r searchHitsCentral-table, message=FALSE, results="asis", echo = FALSE}
```

&nbsp;

(ref:searchHitsWos-caption)  __Web of Science search strategy__ - Full list of search terms used to identify relevant records in the Web of Science bibliographic database. The number of records returned per search term is also displayed.

(ref:searchHitsWos-scaption) Web of Science search strategy

```{r searchHitsWos-table, message=FALSE, results="asis", echo = FALSE}
```

&nbsp;<!----------------------------------------------------------------------->

### Code to search preprints {#appendix-medrxivr-code}

<!-- TODO Double check this is the final code used -->

As detailed extensively in the main text, the inclusion of preprinted literature was a key aspect of the review presented in Chapters \@ref(sys-rev-methods-heading) & \@ref(sys-rev-results-heading). Using the `medrxivr` tool presented in Chapter \@ref(sys-rev-tools-heading).

Given the relatively small number of records contained in the preprint server, no study design filters were employed. For the same reason, the number of search terms is smaller compared with the main database searches, as terms returning no records have been removed.

```{r, eval = FALSE}

library(medrxivr)

# Define exposure search terms
topic1 <- c(mx_caps("statin"),
            mx_caps("ldl"),
            mx_caps("hdl"),
            mx_caps("TG"),
            mx_caps("triglycer"),
            paste0("\\b",mx_caps("TC"),"\\b"),
            mx_caps("ezetim"),
            mx_caps("fibrate"),
            mx_caps("bile acid"),
            mx_caps("lipoprotein"),
            mx_caps("lipid"),
            mx_caps("cholesterol"))

# Define outcome search terms
topic2 <- c(mx_caps("dementia"),
            mx_caps("alzheim"),
            mx_caps("MCI"),
            mx_caps("mild cognitive"))

# Combine search topics using Boolean AND
query <- list(topic1, topic2)

# Run search in bioRxiv
bx_data <- mx_api_content(server = "biorxiv",
                          to_date = "2019-08-01")
bx_results <- mx_search(bx_data, query)

# Run search in medRxiv
mx_data <- mx_api_content(server = "medrxiv",
                          to_date = "2019-08-01")
mx_results <- mx_search(mx_data, query)

```

&nbsp;<!----------------------------------------------------------------------->

### Calculating Gwet's AC1 {#appendix-gwet}

Gwet's agreement coefficient, $AC1$ is defined as:

\begin{equation}
AC1 = \frac{observed\;agreement-chance\;agreement}{1-chance\;agreement}
  (\#eq:AC1-words)
\end{equation}

In reference to a two-by-two table with cells A, B, C and D, it is calculated using the following:

\begin{equation}
  AC1 = \frac{\frac{A+D}{N}-e(\gamma)}{1-e(\gamma)}
  (\#eq:AC1-main)
\end{equation}

where $e(\gamma)$ is the chance agreement between raters, given as $2q(1-q)$, where

\begin{equation}
  q = \frac{(A+C)+(A+B)}{2N}
  (\#eq:AC1-supp)
\end{equation}

&nbsp;<!----------------------------------------------------------------------->

### Risk-of-bias assessment tool for Mendelian randomisation studies {#appendix-mr-rob}

A copy of the risk of bias tool used to assess Mendelian randomisation studies is presented in Table \@ref(tab:mrTool-table), below.

\blandscape

(ref:mrTool-caption) __Mendelian randomisation risk-of-bias assessment tool__ - A copy of the tool used to assess risk of bias in Mendelian randomisation studies, adapted from that developed by Mamluk et al.[@mamluk2020]

(ref:mrTool-scaption) Mendelian randomisation risk-of-bias assessment tool

```{r mrTool-table, message=FALSE, results="asis", echo = FALSE}
```

\elandscape

## Chapter 4

### PRISMA Checklist

\arrayrulecolor{white}

&nbsp;

<!----------------------------------------------------------------------------->
(ref:prismaTab-caption) __PRISMA Checklist__ - Following best practice guidelines, the Section/Table/Figure number where important details on the systematic review can be found are listed.

(ref:prismaTab-scaption) PRISMA Checklist

```{r prismaTab-table, message=FALSE, results="asis", echo = FALSE}
```
<!----------------------------------------------------------------------------->
\arrayrulecolor{black}

&nbsp;

### Funnel plots {#appendix-funnel-plots}

```{r funnelPlots, include = FALSE}
```

<!----------------------------------------------------------------------------->
(ref:funnerlPlots-cap) __Funnel plots__ - To assess the potential for small study effects, funnel plots were produced for the three meta-analyses reported in Chapter \@ref(sys-rev-results-heading) that contained more than 10 results. Visual inspection of the plots did not indicate the presence of small study effect, which was supported in each case by the result of Egger's regression test for funnel plot asymmetry.

(ref:funnerlPlots-scap) Funnel plots

```{r funnerlPlots, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:funnerlPlots-cap)', out.width='100%', fig.scap='(ref:funnerlPlots-scap)'}
knitr::include_graphics(file.path("figures/appendix/funnel_composite.pdf"))
```
<!----------------------------------------------------------------------------->

&nbsp;<!----------------------------------------------------------------------->

### Dose response plots by fraction {#appendix-dr-fraction}

In addition to examining a non-linear relationship between LDL-c and Alzheimer's disease (see Figure \@ref(fig:lipidsDoseResponse)), I performed a dose response analysis for two other exposure-outcome pairs. The results of these analyses are presented below.

(ref:drDemeniaTC-cap) __Dose-response meta-analysis of total cholesterol on all-cause dementia__ - Non-randomised studies examining the association of total cholesterol with all-cause dementia across several categories or "doses" of exposure were synthesised using a dose response meta-analysis. The reference dose was defined _a priori_ as 200 _mg/dL_, the cut-off of the "Normal"/"Optimal" category for total cholesterol as detailed in Table \@ref(tab:lipidLevels-table).

(ref:drDemeniaTC-scap) Dose-response meta-analysis of total cholesterol on all-cause dementia

```{r drDemeniaTC, echo = FALSE, results="asis", fig.align='center', fig.pos = "H", fig.cap='(ref:drDemeniaTC-cap)', out.width='60%', fig.scap='(ref:drDemeniaTC-scap)'}
knitr::include_graphics(file.path("figures/sys-rev/dr_Dementia_TC.png"))
```

(ref:drAdLDL-cap) __Dose-response meta-analysis of low-density lipoprotein cholesterol on Alzheimer's disease__ - Non-randomised studies examining the association of LDL with Alzheimer's disease across several categories or "doses" of exposure were synthesised using a dose response meta-analysis. The reference dose was defined _a priori_ as 100 _mg/dL_, the cut-off of the "Normal"/"Optimal" category for LDL as detailed in Table \@ref(tab:lipidLevels-table).

(ref:drAdLDL-scap) Dose-response meta-analysis of low-density lipoprotein cholesterol on Alzheimer's disease

```{r drAdLDL, echo = FALSE, results="asis", fig.align='center', fig.pos = "H", fig.cap='(ref:drAdLDL-cap)', out.width='60%', fig.scap='(ref:drAdLDL-scap)'}
knitr::include_graphics(file.path("figures/sys-rev/dr_AD_LDL.png"))
```

&nbsp;<!----------------------------------------------------------------------->

## Chapter 5 {#appendix-cprd-analysis}

### Amendments to protocol {#appendix-cprd-amendments}

Any changes made to the approved protocol are "Minor amendments" as per the ISAC criteria, specifically falling under the category of "Additional methods to further control for confounding or sensitivity analysis provided these are to be reported as secondary to the main findings".

Changes include:

-   Use of a time-varying treatment indicator, to correctly classify time-at-risk.
-   Inclusion of additional covariates that are adjusted for in the main model.

### Minimum follow-up period {#cprd-min-fu}

Enforcing a minimum period of follow-up is normally used to reduce the risk of reverse causation in an analysis by excluding events occurring within a certain time-frame post-exposure. With respect to dementia outcomes, which have a long prodromal period, enforcing a minimum period of follow-up can add biological validity to an analysis. Incident cases of dementia that occur within a few years of starting a drug may already be beyond effective preventive treatment at the time the drug was started, and so restricting the analysis to those with longer exposure to treatment may provide a more reliable estimate of the effect.

<!----------------------------------------------------------------------------->
```{r minFollowUp, include = FALSE}
```

(ref:minFU-cap) __Illustration of minimum follow-up period in two study designs__ - Panel A shows a study in which all participants enter the study and are assigned an exposure group at the same time, making it easy to define a minimum period of follow-up. In contrast, Panel B illustrates the time- varying analysis used in Chapter \@ref(cprd-analysis-heading), which makes it challenging to define an unbiased minimum period of follow-up.

(ref:minFU-scap) Illustration of minimum follow-up period in two study designs

```{r minFU, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:minFU-cap)', out.width='100%', fig.scap='(ref:minFU-scap)'}
knitr::include_graphics(file.path("figures/appendix/minFU.png"))
```
<!----------------------------------------------------------------------------->

Panel A of Figure \@ref(fig:minFU) describes a standard analysis of an exposure with a minimum enforced follow-up period (in this case, 10 years). Participants enter the cohort and are classified at that point as either exposed or unexposed, for example, on the basis of a total cholesterol measurement. Participants are followed-up over time, but events occurring before the minimum period of follow-up are discarded.

However, applying a similar approach to the analysis performed in Chapter \@ref(cprd-analysis-heading), illustrated in Panel B of Figure \@ref(fig:minFU), is challenging. The use of a time updating treatment indicator results in two potential starting points for the 10-year lag: entry into the cohort or initiation of statin therapy. As the purpose of this analysis is to investigate the impact of statin treatment, starting the minimum period of follow-up from cohort entry is not helpful as it does not ensure that participants have been exposed to the drug for 10 years. Alternatively, defining the minimum period of follow-up from statin initiation cause issues around the treatment of time-at-risk before statin initiation and around the start point for the follow-up period in those never prescribed a statin. In summary, either approach is likely to be substantially biased.

<!-- With these caveats in mind, below I present a sensitivity analysis of the association of statins with dementia outcomes, enforcing a minimum 10-year period of follow-up from statin initiation in those prescribed a statin.  -->

<!----------------------------------------------------------------------------->
<!-- (ref:cprdMinFu-cap) Sensitivity analysis showing the results of a re-analysis of the CPRD data enforcing a minimum follow-up period of 10 years in the statin group. Events occurring within 10 years of inititating statins were discarded, but the participant-time-at-risk was maintained. No minimum follow-up period was applied in those never taking a statin, due to the difficulties in defining the start date for the follow-up period, as discussed in the text above. -->

<!-- (ref:cprdMinFu-scap) Sensitivity analysis of the CPRD data with a 10 year minimum period of follow-up. -->

<!-- ```{r cprdMinFu, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:cprdMinFu-cap)', out.width='100%', fig.scap='(ref:cprdMinFu-scap)'} -->
<!-- knitr::include_graphics(file.path("figures/hold.png")) -->
<!-- ``` -->
<!----------------------------------------------------------------------------->

&nbsp;<!----------------------------------------------------------------------->

## Chapter 6 {#appendix-ipd-analysis}

<!-- ### Email sent to potental collaborators {#ipd-email-collab} -->

<!-- TODO __To be copied from email__ -->

### Covariate matrix {#appendix-ipd-covariate-matrix}

<!----------------------------------------------------------------------------->
(ref:missingMatrix-caption) __Overview of missing data in the IPD analysis__ - Missing variables/values across cohorts included in the IPD analysis. Missing variables are denoted using a "-", while missing values are described as N (%). Note that LDL is missing completely for the Whitehall and CaPS datasets as this measure was derived from the measures for TC, HDL and TG following multiple imputation of missing values in these variables.

(ref:missingMatrix-scaption) Overview of missing data in the IPD analysis

```{r missingMatrix-table, message=FALSE, results="asis", echo = FALSE}
```
<!----------------------------------------------------------------------------->

\newpage

### Dementia definitions across cohorts {#appendix-ipd-dementia-def}

&nbsp;

<!----------------------------------------------------------------------------->
(ref:ipdDementiaDef-caption) __Definition of dementia outcomes in IPD meta-analysis__ - The criteria used to define dementia outcomes across cohorts included in the individual participant data meta-analysis are listed below. Note the codes used to define dementia and related outcomes in the HES data as part of the Whitehall II cohort were taken from a previously published analysis.[@russ2014]

(ref:ipdDementiaDef-scaption) Definition of dementia outcomes in IPD meta-analysis

```{r ipdDementiaDef-table, message=FALSE, results="asis", echo = FALSE}
```
<!----------------------------------------------------------------------------->

&nbsp;

## Chapter 7 {#appendix-tri-analysis}

### Calculation of average adjustment values from previous expert elicitation exercise {#appendix-adjustment-values}

The paper describing the bias/indirectness-adjusted meta-analysis method used expert elicitation to define a distribution for  bias and indirectness across eight studies examining the effectiveness of routine anti-D prophylaxis in Rhesus negative women.[@turner2009] The authors consider five sources of bias (selection bias, performance bias, attrition bias, detection bias and other bias) and four sources of indirectness (population, intervention, control, outcome). For each domain of bias/indirectness, four experts were asked to define the 67% interval around the effect expected due to bias/indirectness in the absence of a true effect. Additive or multiplicative scales were used to capture the intervals for additive and proportional bias/indirectness, respectively. The authors then derived a mean and variance for each on the log scale. The resulting data for two domains of bias (attrition and detection) in one study are shown in Table \@ref(tab:addTable1-table).

&nbsp;<!----------------------------------------------------------------------->  
(ref:addTable1-caption) __Elicited bias distributions__ -  Distributions of bias, defined by position and variance terms, assigned by four independent experts to attrition and detection bias in a single study. This data was obtained from a previous publication,[@turner2009] and represents an illustrative subset of the larger dataset.

(ref:addTable1-scaption) Elicited bias distributions

```{r addTable1-table, message=FALSE, results="asis", echo = FALSE}
```
&nbsp;<!----------------------------------------------------------------------->

In an attempt to inform the prior distributions used in the triangulation exercise presented in Chapter \@ref(tri-heading), I first averaged the position and variance values across the four experts for each source of bias/indirectness within each study. I took the average of the absolute of the position value, as I wanted to obtain the average adjustment regardless of direction. The results of this process for the example data presented below are shown in Table \@ref(tab:addTable2-table).

&nbsp;<!----------------------------------------------------------------------->  
(ref:addTable2-caption) __Elicited bias distributions, averaged across experts__ - The expert-averaged position and variance values for two example bias domains (attrition and detection) are shown. These values were obtained by averaging the absolute of the values contained in \@ref(tab:addTable3-table) across the four experts.

(ref:addTable2-scaption) Expert-elicited bias distributions, averaged across experts

```{r addTable2-table, message=FALSE, results="asis", echo = FALSE}
```
&nbsp;<!----------------------------------------------------------------------->

I then took the average and maximum of the expert-averaged position and variance values. The results of this for the example data are presented in Table \@ref(tab:addTable3-table). 
&nbsp;<!----------------------------------------------------------------------->  
(ref:addTable3-caption) __Mean and maximum bias distributions, using example data__ - The mean and maximum of position and variance values described in Table \@ref(tab:addTable3-table) were calculated to give the average and maximum bias adjustment across studies and domains of bias.

(ref:addTable3-scaption) Mean and maximum bias distributions, using example data

```{r addTable3-table, message=FALSE, results="asis", echo = FALSE}
```
&nbsp;<!----------------------------------------------------------------------->

This two-step approach (averaging across assessors before taking the average and maximum across domains/studies) was important. While the final average values will be the same whichever approach is taken, in a one-step process, the maximum value can be artificially inflated by a single expert's estimate. For example, in the illustrative data subset above, Assessor 2 assigns a distribution of N(-0.32,0.14), which would double the maximum position (0.16) and triple the maximum variance (0.0475) obtained when first averaging across experts.

The distributions arising from this applying this approach to the entire dataset, stratified by bias/indirectness and additive/proportional effects, are presented in Table \@ref(tab:turnerEstimates-table).

&nbsp;<!----------------------------------------------------------------------->  
(ref:turnerEstimates-caption) __Mean and maximum distributions for bias and indirectness, using full data__ - Using all data resulting from the elicitation exercise reported in _Turner et al_[@turner2009], I calculated mean and maximum distributions for bias and indirectness using the method described above. Note I did not calculate values for additive indirectness, as in my analysis, all indirectness was assumed to be proportional.

(ref:turnerEstimates-scaption) Mean and maximum distributions for bias and indirectness, using full data

```{r turnerEstimates-table, message=FALSE, results="asis", echo = FALSE}
```
&nbsp;<!----------------------------------------------------------------------->

The mean values shown were used to define the distributions assigned to  "Moderate" bias/indirectness judgements in my analysis (see Tables \@ref(tab:priorsAdd-table) & \@ref(tab:priorsIndirect-table)).

&nbsp;<!----------------------------------------------------------------------->

### Comparison of two scenarios of bias

As discussed in the main text, I defined two sets of additive bias distributions as a sensitivity analysis. The results for the two case studies under the two scenarios are presented in the figures below. Under the first, shown on the left of the subsequent  forest plot, the difference in the position value between the distributions assigned to "High" and "Moderate" bias levels is the same as between the "Moderate" and "Low" levels. Under the second scenario, the difference in the position value between the distributions assigned to "High" and "Moderate" is twice that between "Moderate" and "Low".

(ref:fpLdlAdComparison-cap) __Comparison of bias-/indirectness-adjusted meta-analyses of LDL-c/Alzheimer's disease under two different scenarios of additive bias.__ Scenario 1: position value between the additive distributions assigned to "High" and "Moderate" bias levels is the same as between the "Moderate" and "Low" levels. Scenario 2: position value between the additive distributions assigned to "High" and "Moderate" bias levels is twice that between "Moderate" and "Low" levels. See Section \@ref(tri-bias-assign) for more details.

(ref:fpLdlAdComparison-scap) Comparison of bias-/indirectness-adjusted meta-analyses of LDL-c/Alzheimer's disease under two different scenarios of additive bias.
```{r fpLdlAdComparison, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:fpLdlAdComparison-cap)', out.width='100%', fig.scap='(ref:fpLdlAdComparison-scap)'}
knitr::include_graphics(file.path("figures/tri/fp_paired_midlife_ldl_ad_scenarios.png"))
```

(ref:fpTgVaDComparison-cap) __Comparison of bias-/indirectness-adjusted meta-analyses of triglycerides/vascular dementia under two different scenarios of prior distributions of bias.__ Scenario 1: position value between the additive distributions assigned to "High" and "Moderate" bias levels is the same as between the "Moderate" and "Low" levels. Scenario 2: position value between the additive distributions assigned to "High" and "Moderate" bias levels is twice that between "Moderate" and "Low" levels. See Section \@ref(tri-bias-assign) for more details.

(ref:fpTgVaDComparison-scap) Comparison of bias-/indirectness-adjusted meta-analyses of triglycerides/vascular dementia under two different scenarios of prior distributions of bias.

```{r fpTgVaDComparison, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:fpTgVaDComparison-cap)', out.width='100%', fig.scap='(ref:fpTgVaDComparison-scap)'}
knitr::include_graphics(file.path("figures/tri/fp_paired_midlife_tg_vad_scenarios.png"))
```

&nbsp;<!----------------------------------------------------------------------->

## Chapter 8

### Reproducible research {#disc-repro-research}

Reproducible and science has been a key theme running through this thesis, as reflected by the development of an open-source tools to search preprinted literature, create visualisations of risk-of-bias data and to preform quantitative triangulation. In line with this, an open-source copy of the code used to produce this thesis is available on GitHub, as is the code used to perform the analysis contained within it.

Unfortunately, given the nature of the data underpinning the CPRD (Chapter \@ref(cprd-analysis-heading)) and IPD (Chapter \@ref(ipd-heading)) analyses, it is not possible to make it publicly available. Access to the CPRD data is dependent on an application to its managing body, while access to the cohorts within the DPUK requires approval from both the DPUK and individual cohort owners. In the absence of data sharing, the best option to enable review and increase transparency is the open sharing of the code used to perform the analysis.[@goldacre2019]

Finally, all software projects in these thesis attempt to conform to minimal best practices for research computing. [@wilson2014; @wilson2017]

&nbsp;<!----------------------------------------------------------------------->

### Catalogue of failures {#appendix-catalogue-failures}

As part of a reflective process, during the course of my PhD I kept a record of things (analytical mishaps, failed experiments, and unsuccessful grant applications) I attempted but which did not work. These are summarised briefly here:

- Prior to the analysis performed in Chapter \@ref(cprd-analysis-heading), I attempted an analysis using physicians prescribing preference as an instrumental variable to compare different lipid-regulating agent classes for dementia prevention. However, this analysis was not pursued due to the very small number of patients prescribed a lipid-regulating agent other than statins.

- I initially failed to realise that a code list I was using for the positive control outcome of ischemic heart disease (IHD) in the CPRD analysis (Chapter \@ref(cprd-analysis-heading)) included only very severe IHD outcomes. This resulted in the positive protective control analysis giving substantially inflated hazard ratios (HR > 4) and led to substantial confusion in the study team.

- For the IPD analysis in Chapter \@ref(ipd-heading), I had to be reminded that (a) BMI can be calculated from weight and height, and (b) that whether you ask at 45 or 50, the age at which someone left school will be the same.

- I was involved in unsuccessful grant applications to the NIHR (with Andrew Beswick), to the Jean Golding Institute (with James Byrne and Chris Penfold) and to the PGR Hub (with Matthew Lee).

&nbsp;<!----------------------------------------------------------------------->

# Other Appendices {#other-appendix-heading}

\minitoc <!-- this will include a mini table of contents-->

\newpage

## Software used to create this thesis

```{r, include = FALSE}
knitr::write_bib(c(.packages(), "bookdown", "knitr"), here::here("bibliography/packages.bib"))
```

<!-- TODO These citations are not rendering properly - I think because bookdown using this latex format only allows 1 .bib file. -->

This thesis was written in RMarkdown using the RStudio integrated development environment. Several R packages were used as part of this project.`r gen_rmd_citation("packages.bib")` The version number of these packages is shown below:

&nbsp;

```{r, echo = FALSE}
sessionInfo()
```

\newpage

## `robvis`: producing risk-of-bias visualisations {#appendix-robvis}

<!-- TODO Need to complete this sections -->
The `robvis` package provides functions to visualise risk-of-bias assessments, with the resulting plots formatted based on the specific risk-of-bias assessment tool used. The tool is documented extensively online at [https://mcguinlu.github.io/robvis/](https://mcguinlu.github.io/robvis/). As such, it is not described in detail here, other than to illustrate the data structures required for the figures presented in this thesis and to describe its reception in the evidence synthesis community.

&nbsp;<!----------------------------------------------------------------------->

### Data setup

For the paired forest plots presented in Chapter \@ref(sys-rev-results-heading), the tool expects a "Study" column, followed by one column per domain of bias and one column for the overall risk of bias judgement. To illustrate this, the `data_rob2` dataset which is available via the `robvis` package and contains example risk-of-bias assessments performed using the RoB2.0 tool for randomized controlled trials, is shown in Table \@ref(tab:biasData-table).

For the bias direction plots illustrated in Chapter \@ref(tri-heading), additional details on each domain of bias are required by the tool, namely the direction and type of bias in each domain. An example of the data required by `robvis` in order to produce the bias direction plots is shown in \@ref(tab:biasDirectionData-table).

```{r, echo = FALSE, warning=FALSE}

robstats <- cranlogs::cran_downloads("robvis", from = "2019-11-22", to = "2021-12-01")

robvis_downloads <- plyr::round_any(sum(robstats$count), 100, f = floor)

date <- format(as.Date("2021-12-01"),"%B %Y")
```

&nbsp;<!----------------------------------------------------------------------->

### Reception

The tool has been well received by the systematic review community. As of `r date`, the `robvis` R package has been downloaded more than `r medrxivr_downloads` times and the paper introducing the tool had been cited more than 450 times. A copy of this publication is available in Appendix \@ref(published-papers). Development of the tool is ongoing, as evidenced by the new bias direction plot functionality, and there are now a number of contributors to the package.

\blandscape
<!----------------------------------------------------------------------------->
(ref:biasData-caption) __Example `robvis` dataset__ - First five assessments contained in `data_rob2`, an example dataset for the Risk of Bias 2 tool for randomised controlled trials available via the `robvis` package.

(ref:biasData-scaption) Example `robvis` dataset

```{r biasData-table, message=FALSE, results="asis", echo = FALSE}
```
<!----------------------------------------------------------------------------->

&nbsp;

<!----------------------------------------------------------------------------->
(ref:biasDirectionData-caption) __Example bias direction dataset__ - Example risk-of-bias data illustrating the structure required for bias direction plots. $d_ij$, $d_id$, and $d_it$ define the extent (Serious, Moderate, Low), direction (right or left), and type (additive or proportional) of bias in the $i$th domain, respectively Note that, because of the wide nature of the data, I have abbreviated the data in order to display it. A key for interpreting the data can be found below the table.

(ref:biasDirectionData-scaption) Example bias direction dataset

```{r biasDirectionData-table, message=FALSE, results="asis", echo = FALSE}
```
<!----------------------------------------------------------------------------->
\elandscape

\newpage

## Copies of papers arising from this thesis {#published-papers}

The following pages contain copies of the following publications:

- _**McGuinness L. A.**, Higgins J. P. T., Walker, V. M., Davies, N. M., Martin, R. M., Coulthard, E., Davey-Smith, G., Kehoe, P. G., and Ben-Shlomo, Y. (2021) Association of lipid-regulating drugs with dementia and related conditions: an observational study of data from the Clinical Practice Research Datalink. medRxiv. DOI:   [10.1101/2021.10.21.21265131](https://doi.org/10.1101/2021.10.21.21265131)_

- _**McGuinness, L. A.**, and L Schmidt. (2020) medrxivr: Accessing and searching medRxiv and bioRxiv preprint data in R. Journal of Open Source Software 5.54 2651. DOI: [10.21105/joss.02651](https://doi.org/10.21105/joss.02651)_

- _**McGuinness, L. A.**, and Higgins J. P. T. (2020) "Risk‐of‐bias VISualization (robvis): An R package and Shiny web app for visualizing risk‐of‐bias assessments." Research Synthesis Method). DOI: [10.1002/jrsm.1411](https://doi.org/10.1002/jrsm.1411)_

- _**McGuinness, L. A.**, and Sheppard A. L. 2020. “A Descriptive Analysis of the Data Availability Statements Accompanying Medrxiv Preprints and a Comparison with Their Published Counterparts.”  PLOS ONE 16(5): e0250887. DOI: [10.1371/journal.pone.0250887](https://doi.org/10.1371/journal.pone.0250887)_

- _Hennessy, E. A., Acabchuk, R., Arnold, P. A., Dunn, A. G., Foo, Y. Z., Johnson, B. T., Geange, S. R., Haddaway, N. R., Nakagawa, S., Mapanga, W., Mengersen, K., Page, M., Sánchez-Tójar, A. Welch, V.,  **and McGuinness L. A.** (2021). Ensuring Prevention Science Research is Synthesis-Ready for Immediate and Lasting Scientific Impact. Prevention Science . DOI: [10.1007/s11121-021-01279-8](https://doi.org/10.1007/s11121-021-01279-8)_

These manuscripts arose from work performed as part of this thesis, as discussed more fully in Section \@ref(literature-contributions).

<!-- TODO These have been removed so as to not mess-up the plagarism score -->

\includepdf[pages=-,link=true,pagecommand={\thispagestyle{empty}}]{published/cprd_medrxiv.pdf}
\includepdf[pages=-,pagecommand={\thispagestyle{empty}}]{published/medrxivr.pdf}
\includepdf[pages=-,pagecommand={\thispagestyle{empty}}]{published/robvis.pdf}
\includepdf[pages=-,pagecommand={\thispagestyle{empty}}]{published/DAS_comparison.pdf}
\includepdf[pages=-,pagecommand={\thispagestyle{empty}}]{published/prevsci.pdf}