### 2022-04-01 front-and-back-matter/_abstract.Rmd
Range of corrections to abstract

__Background:__ Identification of causal relationships between modifiable risk factors and incidence of dementia is central to the development of evidence-based prevention strategies. [Blood lipids represent such a modifiable risk factor, though considerable uncertainty remains concerning their relationship with dementia outcomes.]{.correction} As such, this thesis sought to triangulate the causal effects of lipid levels (total cholesterol, [low-density-lipoprotein cholesterol (LDL-c), high-density-lipoprotein cholesterol (HDL-c)]{.correction}, and triglycerides) on risk of incident dementia (all-cause dementia, Alzheimer's disease, and vascular dementia) using existing and new evidence.

__Methods:__ Four distinct analyses were conducted. Firstly, a systematic review [and meta-analysis]{.correction} was used to summarise the evidence base. Existing evidence was then supplemented via two primary studies: i) a cohort study of lipid regulating agents and dementia incidence in the Clinical Practice Research Datalink (CPRD); and ii) an individual participant data (IPD) meta-analysis of lipid levels and dementia incidence. Finally, a novel quantitative triangulation framework was proposed, and was then used to integrate the results of the previous three studies.

__Results:__ The systematic review did not identify a consistent effect of blood lipids on any dementia outcome across the study designs considered, though there was some suggestion of a protective effect of LDL-c lowering on all-cause dementia and Alzheimer's disease in observational studies of statin use. Analysis of the CPRD data provided [very]{.correction} weak evidence for [an effect of statins on Alzheimer's disease but suggested a harmful association with the all-cause, vascular and other dementia outcomes,]{.correction} likely due to confounding by indication. The IPD meta-analysis provided some evidence for the association of triglycerides and vascular dementia only [(Odds ratio: 1.25, 95%CI: 1.01-1.54)]{.correction}. Finally, the quantitative triangulation analysis did not provide strong evidence for the causal effect of blood lipids on dementia outcomes.

__Conclusions:__
This thesis provides new evidence concerning the role of blood lipids as a modifiable risk factor for dementia and highlights the uncertainty that still remains in relation to this causal question. In addition, it has developed new evidence synthesis methods and tools.