### 2022-04-07 front-and-back-matter/_abstract.Rmd
Corrections to front and back matter

__Background:__ Identification of causal relationships between modifiable risk factors and incidence of dementia is central to the development of evidence-based prevention strategies. [While blood lipids represent an easily modifiable target for preventative intervention, the evidence for their relationship with dementia outcomes is inconclusive due to the set of contradictory results reported by previous studies]{.correction} As such, this thesis sought to triangulate the causal effects of lipid levels (total cholesterol, [low-density-lipoprotein cholesterol (LDL-c), high-density-lipoprotein cholesterol (HDL-c)]{.correction}, and triglycerides) on risk of incident dementia (all-cause dementia, Alzheimer's disease, and vascular dementia) using existing and new evidence.

__Methods:__ Four distinct analyses were conducted. Firstly, a systematic review [and meta-analysis]{.correction} was used to summarise the evidence base. Existing evidence was then supplemented via two primary studies: i) a cohort study of lipid regulating agents and dementia incidence in the Clinical Practice Research Datalink (CPRD); and ii) an individual participant data (IPD) meta-analysis of lipid levels and dementia incidence. Finally, a novel quantitative triangulation framework was proposed, and was then used to integrate the results of the previous three studies.

__Results:__ The systematic review did not identify a consistent effect of blood lipids on any dementia outcome across the study designs considered, though there was some suggestion of a protective effect of LDL-c lowering on all-cause dementia and Alzheimer's disease in observational studies of statin use. Analysis of the CPRD data provided [very]{.correction} weak evidence for [an effect of statins on Alzheimer's disease but suggested a harmful association with the all-cause, vascular and other dementia outcomes,]{.correction} likely due to confounding by indication. The IPD meta-analysis provided some evidence for the association of triglycerides and vascular dementia only [(odds ratio: 1.25, 95%CI: 1.01-1.54)]{.correction}. Finally, the quantitative triangulation analysis did not provide strong evidence for the causal effect of blood lipids on dementia outcomes.

__Conclusions:__
This thesis provides new evidence concerning the role of blood lipids as a modifiable risk factor for dementia and highlights the uncertainty that still remains in relation to this causal question. In addition, it has developed new evidence synthesis methods and tools.